---
title: "Dissertation - Data Cleaning (Emma)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    toc-depth: 3
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

# Load libraries
library(easystats)
library(patchwork)
library(tidyverse)
```

# 1. Data Loading & Harmonisation

```{r}
# ---- Drop Mint items that were removed in Study 2 final selection ----------
# These clusters were excluded: UrSe (unstable), StaS (vague), SexO (overlaps SexS),
# CaNo (overlaps CaCo), and any attention check column.
drop_unused_mint <- function(df) {
  df |> select(!matches("CaNo|SexO|UrSe|StaS|AttentionCheck", ignore.case = TRUE))
}

# ---- Convert "TRUE"/"FALSE" character columns to 0/1 numeric ---------------
convert_logical_cols <- function(df) {
  df |> mutate(across(
    where(~ is.character(.x) && all(.x %in% c("TRUE", "FALSE", NA_character_))),
    ~ as.integer(.x == "TRUE")
  ))
}

# ---- Align columns between dataframes for merging --------------------------
# Ensures both dataframes have the same columns (fills missing with NA)
align_columns <- function(df1, df2) {
  all_vars <- union(names(df1), names(df2))
  df1[setdiff(all_vars, names(df1))] <- NA
  df2[setdiff(all_vars, names(df2))] <- NA
  list(df1 = df1, df2 = df2)
}
```

# 2. InteroceptionScale Study 1

```{r}
interoception1_data <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/data_participants.csv") |>
  rename(
    # Sexual Arousal Sensitivity (SexS)
    MINT_SexS_9 = Sexual_Genital_Q1,
    MINT_SexS_7 = Sexual_Genital_Q2,
    MINT_SexS_8 = Nociception_Genital_Q1,
    # Cardiorespiratory Confusion (CaCo)
    MINT_CaCo_11 = Nociception_Respiratory_Q2,
    MINT_CaCo_12 = Confusion_Cardiac_Q1,
    MINT_CaCo_10 = Confusion_Respiratory_Q1,
    # Relaxation Awareness (RelA)
    MINT_RelA_5 = Sensitivity_State_Q1,
    MINT_RelA_4 = Sensitivity_State_Q2,
    MINT_RelA_6 = Sensitivity_State_Q3,
    # Cardioception (Card)
    MINT_Card_28 = Sensitivity_Cardiac_Q1,
    MINT_Card_29 = Sensitivity_Cardiac_Q2,
    MINT_Card_30 = Sensitivity_Cardiac_Q3,
    # Respiroception (Resp)
    MINT_Resp_25 = Sensitivity_Respiratory_Q1,
    MINT_Resp_26 = Sensitivity_Respiratory_Q3,
    MINT_Resp_27 = Sensitivity_Respiratory_Q4,
    # Gastroception (Gast)
    MINT_Gast_32 = Sensitivity_Gastric_Q1,
    MINT_Gast_31 = Sensitivity_Gastric_Q2,
    MINT_Gast_33 = Sensitivity_Gastric_Q3,
    # Olfactory Compensation (Olfa)
    MINT_Olfa_23 = Sensitivity_Gastric_Q4,
    MINT_Olfa_24 = Sensitivity_Gastric_Q5,
    MINT_Olfa_22 = Sensitivity_SkinThermo_Q5,
    # Dermal Hypersensitivity (Derm)
    MINT_Derm_16 = Sensitivity_SkinThermo_Q1,
    MINT_Derm_18 = Sensitivity_SkinThermo_Q2,
    MINT_Derm_17 = Sensitivity_SkinThermo_Q7,
    # Expulsion Accuracy (ExAc)
    MINT_ExAc_2 = Accuracy_State_Q2,
    MINT_ExAc_1 = Accuracy_Gastric_Q1,
    MINT_ExAc_3 = Accuracy_Gastric_Q2,
    # Satiety Noticing (Sati)
    MINT_Sati_21 = Accuracy_Gastric_Q4,
    MINT_Sati_19 = Accuracy_Gastric_Q6,
    MINT_Sati_20 = Confusion_Gastric_Q2,
    # Urointestinal Inaccuracy (Urin)
    MINT_Urin_14 = Accuracy_ColonBladder_Q1,
    MINT_Urin_13 = Accuracy_ColonBladder_Q5,
    MINT_Urin_15 = Confusion_ColonBladder_Q1
  )

# Select first 19 columns (demographics/metadata) + all MINT items
interoception1_df <- interoception1_data |>
  select(1:19, contains("MINT_")) |>
  mutate(Source = "InteroceptionScale_Study1")

cat("Interoception Study 1 loaded: N =", nrow(interoception1_df), "\n")
cat("Study 1 MINT items:", sum(str_detect(names(interoception1_df), "^MINT_")), "/ 33\n")
```

Interoception Study 1:  191 participants excluded for failing at least one attention check, and 10 based on measures significantly related to the probability of failing attention checks (namely, the multivariate distance obtained with the OPTICS algorithm, Thériault et al., 2024)

# 3. InteroceptionScale Study 2

```{r}
# N=737. Full 33-item Mint + MAIA, IAS, BPQ, demographics, clinical outcomes.

interoception2_df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study2/data/data_participants.csv") |>
  mutate(Source = "InteroceptionScale_Study2") |>
  rename(
    # Deficit metacluster
    MINT_Urin_13 = MINT_Deficit_Urin_1,
    MINT_Urin_14 = MINT_Deficit_Urin_2,
    MINT_Urin_15 = MINT_Deficit_Urin_3,
    MINT_CaCo_10 = MINT_Deficit_CaCo_4,
    MINT_CaCo_11 = MINT_Deficit_CaCo_5,
    MINT_CaCo_12 = MINT_Deficit_CaCo_6,
    MINT_Olfa_22 = MINT_Deficit_Olfa_10,
    MINT_Olfa_23 = MINT_Deficit_Olfa_11,
    MINT_Olfa_24 = MINT_Deficit_Olfa_12,
    MINT_Sati_19 = MINT_Deficit_Sati_13,
    MINT_Sati_20 = MINT_Deficit_Sati_14,
    MINT_Sati_21 = MINT_Deficit_Sati_15,
    # Awareness metacluster
    MINT_SexS_7 = MINT_Awareness_SexS_19,
    MINT_SexS_9 = MINT_Awareness_SexS_20,
    MINT_SexS_8 = MINT_Awareness_SexS_21,
    MINT_RelA_5 = MINT_Awareness_RelA_28,
    MINT_RelA_4 = MINT_Awareness_RelA_29,
    MINT_RelA_6 = MINT_Awareness_RelA_30,
    MINT_ExAc_3 = MINT_Awareness_ExAc_34,
    MINT_ExAc_1 = MINT_Awareness_ExAc_35,
    MINT_ExAc_2 = MINT_Awareness_ExAc_36,
    # Visceroception metacluster
    MINT_Card_28 = MINT_Sensitivity_Card_37,
    MINT_Card_30 = MINT_Sensitivity_Card_38,
    MINT_Card_29 = MINT_Sensitivity_Card_39,
    MINT_Resp_26 = MINT_Sensitivity_Resp_40,
    MINT_Resp_27 = MINT_Sensitivity_Resp_41,
    MINT_Resp_25 = MINT_Sensitivity_Resp_42,
    MINT_Gast_31 = MINT_Sensitivity_Gast_46,
    MINT_Gast_32 = MINT_Sensitivity_Gast_47,
    MINT_Gast_33 = MINT_Sensitivity_Gast_48,
    MINT_Derm_16 = MINT_Sensitivity_Derm_49,
    MINT_Derm_17 = MINT_Sensitivity_Derm_50,
    MINT_Derm_18 = MINT_Sensitivity_Derm_51
  ) |>
  convert_logical_cols() |>
  drop_unused_mint()  # Remove CaNo, SexO, UrSe, StaS


cat("Interoception Study 2 loaded: N =", nrow(interoception2_df), "\n")
```

Interoception Study 2: 118 excluded for failing at least one attention check and 6 based on multivariate distance (using the same procedure as for study 1). 60 excluded due to missing data from a technical issue.

# 4. FakeArt Dataset

```{r}
# N=359. Mint items already use MINT_Facet_N naming (matching Study 2 after renaming).

fakeart_df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeArt/refs/heads/main/data/rawdata_participants.csv") |>
  mutate(Source = "FakeArt")

# compute and remove function
compute_and_remove <- function(fakeart_df, name="BodyAwareness", pattern=name, method="mean") {
  items <- select(fakeart_df, starts_with(pattern), -contains("AttentionCheck"))
  fakeart_df <- fakeart_df[!names(fakeart_df) %in% names(items)]
  if(method == "mean") {
    fakeart_df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    fakeart_df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  fakeart_df
}

# PHQ-4
fakeart_df <- compute_and_remove(fakeart_df, name="PHQ4_Depression", method="mean")
fakeart_df <- compute_and_remove(fakeart_df, name="PHQ4_Anxiety", method="mean")

## Exclusions 
exclude_fakeart <- list()

# Attention checks

fakeart_checks <- data.frame(
  Participant = fakeart_df$Participant,
  A_MINT = ifelse(fakeart_df$MINT_AttentionCheck == 0, 0, 1),
  A_BAIT = ifelse(fakeart_df$BAIT_AttentionCheck == 6, 0, 1)
)
fakeart_checks$fail <- fakeart_checks$A_MINT + fakeart_checks$A_BAIT
exclude_fakeart$checks <- as.character(fakeart_checks$Participant[fakeart_checks$fail > 0])

fakeart_df <- fakeart_df |>
  filter(!Participant %in% exclude_fakeart$checks)

cat(
  "We excluded", length(exclude_fakeart$checks), "participants (",
  insight::format_percent(length(exclude_fakeart$checks) / nrow(fakeart_df)),
  ") who failed at least one attention check.\n"
)
```

```{r}
cat("FakeArt loaded: N =", nrow(fakeart_df), "\n")
```


# 5. FakeChat Dataset

```{r}
fakechat_df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeChat/refs/heads/main/data/rawdata_participants.csv") |>
  mutate(Source = "FakeChat")

# Store exclusions
exclude_fakechat <- list()

```

## 5.1. Attention Checks

```{r}
checks <- data.frame(
  Participant = fakechat_df$Participant,
  
  A_MINT = ifelse(fakechat_df$MINT_AttentionCheck %in% c(0,1), 0, 1),
  A_BAIT = ifelse(fakechat_df$BAIT_AttentionCheck %in% c(5,6), 0, 1)
)

# count failures
checks$fail <- checks$A_MINT + checks$A_BAIT

# exclude anyone failing at least one
exclude_fakechat$checks <- as.character(
  checks$Participant[checks$fail > 0]
)

# exclude anyone failing both attention checks
#exclude_fakechat$checks <- as.character(
#  checks$Participant[checks$fail == 2]
#)

cat(
  "We excluded", length(exclude_fakechat$checks), "participants (",
  insight::format_percent(length(exclude_fakechat$checks) / nrow(fakechat_df)),
  ") who failed at least one attention check.\n"
)

#cat(
#  "We excluded", length(exclude_fakechat$checks), "participants (",
#  insight::format_percent(length(exclude_fakechat$checks) / nrow(fakechat_df)),
#  ") who failed both of the attention checks.\n"
#)
```

## 5.2. Duration Exclusion

```{r}
exclude_fakechat$duration <- as.character(
  fakechat_df$Participant[fakechat_df$Experiment_Duration < 5]
)

exclude_fakechat$duration <- exclude_fakechat$duration[
  !exclude_fakechat$duration %in% exclude_fakechat$checks
]

cat(
  "We excluded", length(exclude_fakechat$duration), "participants (",
  insight::format_percent(length(exclude_fakechat$duration) / nrow(fakechat_df)),
  ") for having completed the experiment in less than 5 minutes.\n"
)
```

## 5.3. Low Quality Exclusion

```{r}
exclude_fakechat$feedback <- as.character(
  fakechat_df$Participant[fakechat_df$Feedback_Quality %in% c(0,1)]
)

exclude_fakechat$feedback <- exclude_fakechat$feedback[
  !exclude_fakechat$feedback %in% c(exclude_fakechat$checks, exclude_fakechat$duration)
]

cat(
  "We excluded", length(exclude_fakechat$feedback), "participants (",
  insight::format_percent(length(exclude_fakechat$feedback) / nrow(fakechat_df)),
  ") for self-reporting that they did not do the experiment carefully and thoroughly.\n"
)
```

## 5.4. Remove Excluded Participants

```{r}
fakechat_df <- fakechat_df |>
  filter(!Participant %in% c(exclude_fakechat$checks, exclude_fakechat$duration, exclude_fakechat$feedback))

cat("Participants after exclusions: N =", nrow(fakechat_df), "\n")
```

## 5.5. Compute Questionnaire Scores

```{r}
# compute and remove function
compute_and_remove <- function(fakechat_df, name, pattern=name, method="mean") {
  items <- select(fakechat_df, starts_with(pattern), -contains("AttentionCheck"))
  fakechat_df <- fakechat_df[!names(fakechat_df) %in% names(items)]
  if(method == "mean") {
    fakechat_df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    fakechat_df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  fakechat_df
}

# PHQ-4
fakechat_df <- compute_and_remove(fakechat_df, name="PHQ4_Depression", method="mean")
fakechat_df <- compute_and_remove(fakechat_df, name="PHQ4_Anxiety", method="mean")
```

## 5.6. Mental Health

```{r}
fakechat_df$Disorders_Psychiatric_Mood <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)
fakechat_df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  fakechat_df$Disorders_Psychiatric_Mood & !is.na(fakechat_df$Disorders_PsychiatricTreatment) & str_detect(fakechat_df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)

fakechat_df$Disorders_Psychiatric_Anxiety <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "Panic|GAD|Social Phobia|Phobia"), TRUE, FALSE)
fakechat_df$Disorders_Psychiatric_AnxietyTreatment <- ifelse(
  fakechat_df$Disorders_Psychiatric_Anxiety & !is.na(fakechat_df$Disorders_PsychiatricTreatment) & str_detect(fakechat_df$Disorders_PsychiatricTreatment, "Anxiolytic|Psychotherapy|Mindfulness"), TRUE, FALSE)

fakechat_df$Disorders_Psychiatric_Eating <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "Eating"), TRUE, FALSE)
fakechat_df$Disorders_Psychiatric_Addiction <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "Addiction"), TRUE, FALSE)
fakechat_df$Disorders_Psychiatric_Borderline <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "BPD"), TRUE, FALSE)

fakechat_df$Disorders_Psychiatric_Autism <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "ASD"), TRUE, FALSE)
fakechat_df$Disorders_Psychiatric_ADHD <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "ADHD"), TRUE, FALSE)


# fakechat_df$Disorders_Psychiatric_Other <- ifelse(str_detect(fakechat_df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)
```

## 5.7. Somatic Health

```{r}
somatic <- data.frame(
  Hypermobility = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Hypermobility Syndrome"), TRUE, FALSE),
  Fibromyalgia = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Fibromyalgia"), TRUE, FALSE),
  ChronicFatigue = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Chronic Fatigue Syndrome"), TRUE, FALSE),
  ChronicPain = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Chronic Pain Syndrome"), TRUE, FALSE),
  BackPain = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Back Pain"), TRUE, FALSE),
  MuscleTension = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Muscle Tension"), TRUE, FALSE),
  SkinRashes = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Skin Rashes"), TRUE, FALSE),
  Eczema = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Eczema"), TRUE, FALSE),
  Psoriasis = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Psoriasis"), TRUE, FALSE),
   # Dermatological = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Other Dermatological"), TRUE, FALSE),
  # Musculoskeletal = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Other Musculoskeletal"), TRUE, FALSE),
  Sjogrens = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Sjogren"), TRUE, FALSE),
  ChestPain = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Chest Pain"), TRUE, FALSE),
  Arrhythmia = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Cardiac Arrhythmia"), TRUE, FALSE),
  Hypertension = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Hypertension"), TRUE, FALSE),
  Hypotension = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Hypotension"), TRUE, FALSE),
  IBS = ifelse(str_detect(fakechat_df$Disorders_Somatic, "IBS"), TRUE, FALSE),
  GERD = ifelse(str_detect(fakechat_df$Disorders_Somatic, "GERD"), TRUE, FALSE),
  Crohn = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Crohn"), TRUE, FALSE),
  UlcerativeColitis = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Ulcerative Colitis"), TRUE, FALSE),
  Celiac = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Celiac Disease"), TRUE, FALSE),
  Gluten = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Gluten"), TRUE, FALSE),
  Lactose = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Lactose"), TRUE, FALSE),
  ShortBreath = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Shortness of Breath"), TRUE, FALSE),
  Asthma = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Asthma"), TRUE, FALSE),
  COPD = ifelse(str_detect(fakechat_df$Disorders_Somatic, "COPD"), TRUE, FALSE),
  SleepApnea = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Sleep Apnea"), TRUE, FALSE),
  Bronchitis = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Chronic Bronchitis"), TRUE, FALSE),
  Nausea = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Nausea/Vomiting"), TRUE, FALSE),
  Dizziness = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Dizziness/Lightheadedness"), TRUE, FALSE),
  Migraine = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Migraine"), TRUE, FALSE),
  Neuropathy = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Neuropathy"), TRUE, FALSE),
  Epilepsy = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Epilepsy"), TRUE, FALSE),
  MS = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Multiple Sclerosis"), TRUE, FALSE),
  FrequentUrination = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Frequent Urination"), TRUE, FALSE),
  Endometriosis = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Endometriosis"), TRUE, FALSE),
  Cystitis = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Interstitial Cystitis"), TRUE, FALSE),
  PelvicPain = ifelse(str_detect(fakechat_df$Disorders_Somatic, "Chronic Pelvic Pain Syndrome"), TRUE, FALSE)
)


# Cluster A: Afferent Sensitivity (neurogenic excitability, interoceptive hypervigilance)
fakechat_df$Disorders_Somatic_AfferentSensitivity_N <- rowSums(somatic[, c(
  "Migraine", "Neuropathy", "MuscleTension", "Dizziness", 
  "Nausea", "Epilepsy", "FrequentUrination"
)])
fakechat_df$Disorders_Somatic_AfferentSensitivity <- fakechat_df$Disorders_Somatic_AfferentSensitivity_N >= 1

# Cluster B: Central Sensitization (pain and fatigue syndromes)
fakechat_df$Disorders_Somatic_CentralSensitization_N <- rowSums(somatic[, c(
  "Fibromyalgia", "ChronicPain", "ChronicFatigue", "BackPain", 
  "PelvicPain", "Endometriosis", "IBS", "Cystitis"
)])
fakechat_df$Disorders_Somatic_CentralSensitization <- fakechat_df$Disorders_Somatic_CentralSensitization_N >= 1

# Cluster C: Autonomic Dysfunction (dysautonomia and cardiorespiratory instability)
fakechat_df$Disorders_Somatic_AutonomicDysfunction_N <- rowSums(somatic[, c(
  "Hypermobility", "Arrhythmia", "ChestPain", "ShortBreath", 
  "Hypotension", "Hypertension", "SleepApnea", "COPD", "Bronchitis"
)])
fakechat_df$Disorders_Somatic_AutonomicDysfunction <- fakechat_df$Disorders_Somatic_AutonomicDysfunction_N >= 1

# Cluster D: Immune-Inflammatory (gut-brain axis, autoimmune and barrier-related conditions)
fakechat_df$Disorders_Somatic_ImmuneInflammatory_N <- rowSums(somatic[, c(
  "SkinRashes", "Eczema", "Psoriasis", "Sjogrens", "Asthma", 
  "Celiac", "Gluten", "Lactose", "Crohn", "UlcerativeColitis", 
  "GERD", "MS"
)])
fakechat_df$Disorders_Somatic_ImmuneInflammatory <- fakechat_df$Disorders_Somatic_ImmuneInflammatory_N >= 1
```

## 5.8. FakeChat Final Sample

```{r}
cat("FakeChat loaded: N =", nrow(fakechat_df), "\n")
```

FakeChat: 58 participants excluded for failing at least one attention check.

 
# 6. Merge All Datasets

```{r}
# combine all datasets
datasets <- list(
  interoception1_df,
  interoception2_df,
  fakeart_df,
  fakechat_df
)

mega_df <- bind_rows(datasets) |>
  select(
    "Participant", "Gender", "Age", "Education", "Discipline", "Student", "Ethnicity", "Country", "Source", "LifeSatisfaction", "Disorders_Psychiatric", "Disorders_PsychiatricTreatment", "Disorders_Somatic", "PHQ4_Anxiety", "PHQ4_Depression", "Disorders_Psychiatric_Mood", "Disorders_Psychiatric_MoodTreatment", "Disorders_Psychiatric_Anxiety", "Disorders_Psychiatric_AnxietyTreatment", "Disorders_Psychiatric_Eating", "Disorders_Psychiatric_Addiction", "Disorders_Psychiatric_Borderline", "Disorders_Psychiatric_Autism", "Disorders_Psychiatric_ADHD", "Disorders_Somatic_AfferentSensitivity", "Disorders_Somatic_AfferentSensitivity_N", "Disorders_Somatic_CentralSensitization", "Disorders_Somatic_CentralSensitization_N", "Disorders_Somatic_AutonomicDysfunction", "Disorders_Somatic_AutonomicDysfunction_N", "Disorders_Somatic_ImmuneInflammatory", "Disorders_Somatic_ImmuneInflammatory_N", starts_with("MINT_")
  )

mega_df$Gender[mega_df$Gender == "other"] <- "Other"

mega_df <- mega_df |> 
  filter(Age >= 18)

# datasets <- append(datasets, list(new_dataset))   ## use to add more data if necessary

cat("\n=== FINAL MERGED DATASET ===\n")
cat("Total N =", nrow(mega_df), "\n")

cat("Sources:\n")
print(table(mega_df$Source, useNA = "ifany"))

# Identify MINT items
mint_items <- select(mega_df, starts_with("MINT_")) |>
  select(-MINT_AttentionCheck)

cat("\nMINT items in merged dataset:", length(mint_items), "\n")

# Missingness check by dataset
#mega_df |>
#  group_by(Source) |>
#  summarise(
#    n = n(),
#    pct_missing_mint = round(mean(is.na(across(all_of(mint_items)))), 3),
#    .groups = "drop"
#  ) |>
#  print()
```


# 7. Final Sample

::: {.panel-tabset}


```{r}
cat("The final sample includes", report_participants(mega_df, threshold = 5, age = "Age", gender = "Gender", race = "Ethnicity"))
```

## Age and Gender

```{r p_demographics_age_gender}
p_age <- mega_df |>
  ggplot(aes(x = Age, fill = Gender)) + 
  geom_histogram(data=mega_df, aes(x = Age, fill=Gender), binwidth = 2) +
  geom_vline(xintercept = mean(mega_df$Age), color = "red", linewidth=1.5, linetype="dashed") +
  scale_fill_viridis_d() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(20, max(mega_df$Age), by = 10 )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

p_age
```

## Education

```{r p_demographics_education}
#| warning: false
#| error: false

p_edu <- mega_df |>
  mutate(Student = ifelse(is.na(Student), FALSE, Student),
         Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Student)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level", fill = "Is currently a student") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu
```

## Ethnicity

```{r p_demographics_ethnicity}
#| warning: false
#| error: false

p_eth <- mega_df |>
  filter(!is.na(Ethnicity)) |> 
  ggplot(aes(x = "", fill = Ethnicity)) +
  geom_bar() +
  coord_polar("y") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )
    
p_eth
```

## Mental Health Conditions

```{r}
#| warning: false
#| error: false

select(mega_df, Participant, Gender, 
             Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment,
             Disorders_Psychiatric_Anxiety, Disorders_Psychiatric_AnxietyTreatment,
             Disorders_Psychiatric_Eating, Disorders_Psychiatric_Addiction, Disorders_Psychiatric_Borderline,
             Disorders_Psychiatric_Autism, Disorders_Psychiatric_ADHD) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "AnxietyTreatment", "Anxiety (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value, na.rm = TRUE) / sum(!is.na(Value)), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot)),
         Treatment = ifelse(str_detect(Disorder,  "Treatment"), "With treatment", "Without treatment")) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

## Somatic Health Conditions

```{r}
somatic |> 
  select(where(~n_distinct(.) > 1)) |> 
  correlation::correlation(redundant = TRUE, p_adjust="none") |>
  correlation::cor_sort() |> 
  mutate(label = ifelse(p < .001, "*", ""),
         label = ifelse(r == 1, "", label)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  geom_text(aes(label=label), size=3) +
  scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
  labs(title = "Co-occurence Matrix of Somatic Disorders") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_blank())

somatic2 <- mega_df |> 
  select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
  select(ends_with("_N"), Gender, Participant) |>
  pivot_longer(starts_with("Disorders_Somatic_")) |>
  filter(!is.na(value)) |>
  summarize(n = n(), .by=c("name", "value", "Gender")) |> 
  mutate(name = str_remove(name, "Disorders_Somatic_"),
         name = str_remove(name, "_N")) |> 
  ggplot(aes(x = value, y=n, fill=Gender)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Number of Issues") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~name)
somatic2
```


:::


# 8. Save data

```{r}
write.csv(mega_df, "C:/Users/emmal/OneDrive - University of Sussex/GitHub/Dissertation/ChatExtracts/data/Emma_dissertationdata.csv", row.names = FALSE)
```
