---
title: "InteroceptionProfiles - Data Analysis (Emma)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    toc-depth: 3
execute: 
  cache: true
---

## 1. Load packages

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(cluster)
library(factoextra)
library(mclust)
library(pvclust)
library(NbClust)
library(dbscan)
library(kernlab)
library(apcluster)
library(nnet)
library(umap)
library(Rtsne)
library(EGAnet)
library(ggdist)

set.seed(42)
```


## 2. Compute Facet & Metacluster Scores

```{r}
# Lower order facet scores (mean of 3 items each)
mega_df <- mega_df |>
  mutate(
    # Awareness metacluster
    MINT_SexS  = rowMeans(across(starts_with("MINT_SexS_")),  na.rm = TRUE),
    MINT_ExAc  = rowMeans(across(starts_with("MINT_ExAc_")),  na.rm = TRUE),
    MINT_RelA  = rowMeans(across(starts_with("MINT_RelA_")),  na.rm = TRUE),
    # Deficit metacluster
    MINT_CaCo  = rowMeans(across(starts_with("MINT_CaCo_")),  na.rm = TRUE),
    MINT_Urin  = rowMeans(across(starts_with("MINT_Urin_")),  na.rm = TRUE),
    MINT_Derm  = rowMeans(across(starts_with("MINT_Derm_")),  na.rm = TRUE),
    MINT_Sati  = rowMeans(across(starts_with("MINT_Sati_")),  na.rm = TRUE),
    MINT_Olfa  = rowMeans(across(starts_with("MINT_Olfa_")),  na.rm = TRUE),
    # Visceroception metacluster
    MINT_Card  = rowMeans(across(starts_with("MINT_Card_")),  na.rm = TRUE),
    MINT_Resp  = rowMeans(across(starts_with("MINT_Resp_")),  na.rm = TRUE),
    MINT_Gast  = rowMeans(across(starts_with("MINT_Gast_")),  na.rm = TRUE)
  )

# Higher order metacluster scores (mean of facets)
mega_df <- mega_df |>
  mutate(
    MINT_Awareness = rowMeans(across(c(MINT_SexS, MINT_ExAc, MINT_RelA)), na.rm = TRUE),
    MINT_Deficit = rowMeans(across(c(MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa)), na.rm = TRUE),
    MINT_Visceroception = rowMeans(across(c(MINT_Card, MINT_Resp, MINT_Gast)), na.rm = TRUE)
  )

mint_lowerorder <- mega_df |> 
  select(MINT_SexS, MINT_ExAc, MINT_RelA, MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa, MINT_Card, MINT_Resp, MINT_Gast)

mint_higherorder <- mega_df |> 
  select(MINT_Awareness, MINT_Deficit, MINT_Visceroception)

cat("Facet scores computed for N =", nrow(mega_df), "participants\n")
cat("Sources included:", paste(unique(mega_df$Source), collapse = ", "), "\n")

# Structure Validation
names(mint_items) <- str_remove(names(mint_items), "MINT_")
```

### 2b. Item Labels

```{r}
labels <- list(
  MINT_ExAc_1 = 
    "I can always accurately feel when I am about to fart",
  MINT_ExAc_2 = 
    "I can always accurately feel when I am about to sneeze",
  MINT_ExAc_3 = 
    "I can always accurately feel when I am about to burp",
  MINT_RelA_4 = 
    "I always feel in my body if I am relaxed",
  MINT_RelA_5 = 
    "I always know when I am relaxed",
  MINT_RelA_6 =
    "My body is always in the same specific state when I am relaxed",
  MINT_SexS_7 =
    "During sex or masturbation, I often feel very strong sensations coming from my genital areas",
  MINT_SexS_8 =
    "My genital organs are very sensitive to pleasant stimulations",
  MINT_SexS_9 =
    "When I am sexually aroused, I often notice specific sensations in my genital area (e.g., tingling, warmth, wetness, stiffness, pulsations)",
  MINT_CaCo_10 =
    "Sometimes my breathing becomes erratic or shallow and I often don't know why",
  MINT_CaCo_11 =
    "I often feel like I can't get enough oxygen by breathing normally",
  MINT_CaCo_12 =
    "Sometimes my heart starts racing and I often don't know why",
  MINT_Urin_13 =
    "I sometimes feel like I need to urinate or defecate but when I go to the bathroom I produce less than I expected",
  MINT_Urin_14 =
    "I often feel the need to urinate even when my bladder is not full",
  MINT_Urin_15 =
    "Sometimes I am not sure whether I need to go to the toilet or not (to urinate or defecate)",
  MINT_Derm_16 = 
    "In general, my skin is very sensitive",
  MINT_Derm_17 = 
    "My skin is susceptible to itchy fabrics and materials",
  MINT_Derm_18 =
    "I can notice even very subtle stimulations to my skin (e.g., very light touches)",
  MINT_Sati_19 =
    "I don't always feel the need to eat until I am really hungry",
  MINT_Sati_20 =
    "Sometimes I don't realise I was hungry until I ate something ",
  MINT_Sati_21 =
    "I don't always feel the need to drink until I am really thirsty",
  MINT_Olfa_22 = 
    "I often check the smell of my armpits",
  MINT_Olfa_23 = 
    "I often check the smell of my own breath",
  MINT_Olfa_24 = 
    "I often check the smell of my farts",
  MINT_Resp_25 = 
    "In general, I am very sensitive to changes in my breathing",
  MINT_Resp_26 = 
    "I can notice even very subtle changes in my breathing",
  MINT_Resp_27 =
    "I am always very aware of how I am breathing, even when I am calm",
  MINT_Card_28 = 
    "In general, I am very sensitive to changes in my heart rate",
  MINT_Card_29 = 
    "I often notice changes in my heart rate",
  MINT_Card_30 =
    "I can notice even very subtle changes in the way my heart beats",
  MINT_Gast_31 =
    "I can notice even very subtle changes in what my stomach is doing",
  MINT_Gast_32 = 
    "In general, I am very sensitive to what my stomach is doing",
  MINT_Gast_33 =
    "I am always very aware of what my stomach is doing, even when I am calm"
)
```

```{r}
cleanlabels <- function(x, qname=FALSE) {
  is_factor <- FALSE
  if(is.factor(x)) {
    is_factor <- TRUE
    vec <- x
    x <- levels(x)
  }
  
  x <- x |> 
    str_replace("MINT_", ifelse(qname, "MINT - ", "")) |> 
    str_replace("_", " - ") 
  
  if(is_factor) {
    levels(vec) <- x
    return(vec)
  }
  x
}
```


## 3. Replicate EGA from InteroceptionScale Study 2

```{r}
#| eval: false

ega <- bootEGA(
  mint_items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  # consensus.method = "lowest_tefi",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, 
  ncores = 4)

save(ega, file="models/ega.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega.RData")

EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega, IS.plot=FALSE)
p_stab <- plot(itemstability)
p_stab

EGAnet::genTEFI(ega$EGA)


ega$EGA
```

Lower-order dimensions are very stable. Higher-order dimensions are mostly stable but a few items show ambiguity (Derm 16, 17 & 18, Olfa 22, 23 & 24).

Deficit metacluster is less consistent (0.546) - less robust, possibly due to heterogeneous items

Network characteristics:
Lower-order network - 33 nodes, 194 edges, edge density 0.367.
Most edges are small (0.071 ± 0.108), range -0.088 to 0.48 → mostly weak to moderate partial correlations.
Higher-order network - 11 nodes (representing lower-order communities), 36 edges, edge density 0.655 → more connected than lower-order network.
Edge weights are slightly stronger on average (0.096 ± 0.103), some negative → stronger inter-community relationships, but some inverse relationships.

Total Entropy Fit Index:
More negative TEFI = better fit of bifactor/hierarchical model.
Results suggest a bifactor structure fits better than just correlated lower-order factors.


## 4. Transpose data to make participants into columns

```{r}
#transposed_mint_lowerorder <- as.data.frame(t(mint_lowerorder))
```


## 5. Clustering

```{r}
# Scale data

mint_scaled <- as.data.frame(scale(mint_lowerorder))

```


### 5a. Estimate Number of Clusters

```{r}
cluster_estimate <- n_clusters(
  mint_scaled,
  package = c("all"),
  n_max = 10,
)
cluster_estimate

plot(cluster_estimate)
```

Method Agreement Procedure:
The choice of 2 clusters is supported by 9 (31.03%) methods out of 29 (Elbow, Silhouette, Ch, CCC, DB, Duda, Pseudot2, Beale, Mcclain).

3 clusters also appears to be equally supported.

Cluster estimation suggested 2–3 clusters. While a 2-cluster solution captured a general low vs high distinction, a 3-cluster solution revealed an additional subgroup characterised by globally reduced interoceptive sensitivity. Given its interpretability and theoretical relevance, the 3-cluster solution was retained.

### 5b. Kmeans

```{r}
# 2 clusters

rez_kmeans2 <- cluster_analysis(mint_scaled, n = 2, method = "kmeans")

rez_kmeans2

plot(summary(rez_kmeans2)) # Visualize cluster centers

predict(rez_kmeans2) # Get clusters


# 3 clusters

rez_kmeans3 <- cluster_analysis(mint_scaled, n = 3, method = "kmeans")

rez_kmeans3

plot(summary(rez_kmeans3)) # Visualize cluster centers

predict(rez_kmeans3) # Get clusters
```

Cluster 1 (Low across most facets) - Globally low interoception
Especially low on:
Cardioception (-0.84)
Respiroception (-0.82)
Dermal Hypersensitivity (-0.58)

Cluster 2 - High awareness/relatively low confusion
High on: Awareness facets (SexS, ExAc, RelA) and Visceroception (Card, Resp, Gast)
But slightly low on deficit facets (CaCo, Urin)

Cluster 3 (Mixed pattern) - High confusion/dysregulated interoception
High on Deficit facets (CaCo = +0.94, Urin = +0.71)
Moderate/high elsewhere


### 5c. Hierarchical Clustering

```{r}
rez_hclust <- cluster_analysis(mint_scaled, n = 3, method = "hclust")

rez_hclust

# Visualize
plot(rez_hclust) + theme_modern()

predict(rez_hclust)
```

### 5d. Hierarchical K-Means

```{r}
rez_hkmeans <- cluster_analysis(mint_scaled, n = 3, method = "hkmeans")

rez_hkmeans
```

### 5e. K-Medoids (PAM)

```{r}
rez_pam <- cluster_analysis(mint_scaled, n = 3, method = "pam")

rez_pam

# Visualize
plot(rez_pam) + theme_modern()
```

### 5f. Bootstrapped Hierarchical Clustering

```{r}
# WARNING - SLOW

#rez_hclust2 <- cluster_analysis(mint_scaled,
#  n = NULL,
#  method = "hclust",
#  iterations = 500,
#  ci = 0.90
#)

#rez_hclust2
#plot(rez_hclust2) + theme_modern() # Visualize
```

Suggests the existence of 0 clusters.

### 5g. DBSCAN - REMOVE?

```{r}
eps <- n_clusters_dbscan(mint_scaled, min_size = 0.1)
eps
plot(eps) # Find the 'elbow' of the curve: 4.25
```

```{r}
rez_dbscan <- cluster_analysis(mint_scaled, method = "dbscan", dbscan_eps = 4.25)

rez_dbscan
plot(rez_dbscan) + theme_modern() # Visualize
```

Suggests the existence of 1 cluster.
The 2 clusters accounted for 1.47% of the total variance of the original data.

### 5h. Hierarchical DBSCAN - REMOVE?

```{r}
rez_hdbscan <- cluster_analysis(mint_scaled, method = "hdbscan")

rez_hdbscan

# Visualize
plot(rez_hdbscan) + theme_modern()
```

The unique cluster accounted for 0.00% of the total variance of the original data.
Suggests the existance of 0 clusters.

### 5i. K-Medoids with estimation of number of clusters (pamk)

```{r}
rez_pamk <- cluster_analysis(mint_scaled, method = "pamk")

rez_pamk

# Visualize
plot(rez_pamk) + theme_modern()
```

The 2 clusters accounted for 17.50% of the total variance of the original data.

### 5j. Mixture - REMOVE?

```{r}
rez_mixture <- cluster_analysis(mint_scaled, method = "mixture")

rez_mixture

# Visualize
plot(rez_mixture) + theme_modern()
```

The 4 clusters accounted for 13.43% of the total variance of the original data.

## 6. Metaclustering

```{r}
list_of_results <- list(
  rez_kmeans3, rez_hclust, rez_hkmeans, rez_pam, rez_pamk
)

probability_matrix <- cluster_meta(list_of_results)

# Plot the matrix as a reordered heatmap
heatmap(unclass(probability_matrix),
  scale = "none",
  col = grDevices::hcl.colors(256, palette = "inferno")
)
```

