---
title: "InteroceptionProfiles - Data Analysis (Emma)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    toc-depth: 3
execute: 
  cache: true
---

## 0. Load packages

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(easystats)
library(cluster)
library(factoextra)
library(mclust)
library(pvclust)
library(NbClust)
library(dbscan)
library(kernlab)
library(apcluster)
library(nnet)
library(umap)
library(Rtsne)
library(EGAnet)
library(ggdist)

set.seed(42)
```


## 1. Data Loading & Harmonisation

```{r}
# ---- Drop Mint items that were removed in Study 2 final selection ----------
# These clusters were excluded: UrSe (unstable), StaS (vague), SexO (overlaps SexS),
# CaNo (overlaps CaCo), and any attention check column.
drop_unused_mint <- function(df) {
  df |> select(!matches("CaNo|SexO|UrSe|StaS|AttentionCheck", ignore.case = TRUE))
}

# ---- Convert "TRUE"/"FALSE" character columns to 0/1 numeric ---------------
convert_logical_cols <- function(df) {
  df |> mutate(across(
    where(~ is.character(.x) && all(.x %in% c("TRUE", "FALSE", NA_character_))),
    ~ as.integer(.x == "TRUE")
  ))
}

# ---- Align columns between dataframes for merging --------------------------
# Ensures both dataframes have the same columns (fills missing with NA)
align_columns <- function(df1, df2) {
  all_vars <- union(names(df1), names(df2))
  df1[setdiff(all_vars, names(df1))] <- NA
  df2[setdiff(all_vars, names(df2))] <- NA
  list(df1 = df1, df2 = df2)
}
```

### 1b. InteroceptionScale Study 1

```{r}
interoception1_data <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study1/data/data_participants.csv") |>
  rename(
    # Sexual Arousal Sensitivity (SexS)
    MINT_SexS_9 = Sexual_Genital_Q1,
    MINT_SexS_7 = Sexual_Genital_Q2,
    MINT_SexS_8 = Nociception_Genital_Q1,
    # Cardiorespiratory Confusion (CaCo)
    MINT_CaCo_11 = Nociception_Respiratory_Q2,
    MINT_CaCo_12 = Confusion_Cardiac_Q1,
    MINT_CaCo_10 = Confusion_Respiratory_Q1,
    # Relaxation Awareness (RelA)
    MINT_RelA_5 = Sensitivity_State_Q1,
    MINT_RelA_4 = Sensitivity_State_Q2,
    MINT_RelA_6 = Sensitivity_State_Q3,
    # Cardioception (Card)
    MINT_Card_28 = Sensitivity_Cardiac_Q1,
    MINT_Card_29 = Sensitivity_Cardiac_Q2,
    MINT_Card_30 = Sensitivity_Cardiac_Q3,
    # Respiroception (Resp)
    MINT_Resp_25 = Sensitivity_Respiratory_Q1,
    MINT_Resp_26 = Sensitivity_Respiratory_Q3,
    MINT_Resp_27 = Sensitivity_Respiratory_Q4,
    # Gastroception (Gast)
    MINT_Gast_32 = Sensitivity_Gastric_Q1,
    MINT_Gast_31 = Sensitivity_Gastric_Q2,
    MINT_Gast_33 = Sensitivity_Gastric_Q3,
    # Olfactory Compensation (Olfa)
    MINT_Olfa_23 = Sensitivity_Gastric_Q4,
    MINT_Olfa_24 = Sensitivity_Gastric_Q5,
    MINT_Olfa_22 = Sensitivity_SkinThermo_Q5,
    # Dermal Hypersensitivity (Derm)
    MINT_Derm_16 = Sensitivity_SkinThermo_Q1,
    MINT_Derm_18 = Sensitivity_SkinThermo_Q2,
    MINT_Derm_17 = Sensitivity_SkinThermo_Q7,
    # Expulsion Accuracy (ExAc)
    MINT_ExAc_2 = Accuracy_State_Q2,
    MINT_ExAc_1 = Accuracy_Gastric_Q1,
    MINT_ExAc_3 = Accuracy_Gastric_Q2,
    # Satiety Noticing (Sati)
    MINT_Sati_21 = Accuracy_Gastric_Q4,
    MINT_Sati_19 = Accuracy_Gastric_Q6,
    MINT_Sati_20 = Confusion_Gastric_Q2,
    # Urointestinal Inaccuracy (Urin)
    MINT_Urin_14 = Accuracy_ColonBladder_Q1,
    MINT_Urin_13 = Accuracy_ColonBladder_Q5,
    MINT_Urin_15 = Confusion_ColonBladder_Q1
  )

# Select first 19 columns (demographics/metadata) + all MINT items
interoception1_df <- interoception1_data |>
  select(1:19, contains("MINT_")) |>
  mutate(Source = "InteroceptionScale_Study1")

cat("Interoception Study 1 loaded: N =", nrow(interoception1_df), "\n")
cat("Study 1 MINT items:", sum(str_detect(names(interoception1_df), "^MINT_")), "/ 33\n")
```

Interoception Study 1:  191 participants excluded for failing at least one attention check, and 10 based on measures significantly related to the probability of failing attention checks (namely, the multivariate distance obtained with the OPTICS algorithm, Thériault et al., 2024)

### 1c. InteroceptionScale Study 2

```{r}
# N=737. Full 33-item Mint + MAIA, IAS, BPQ, demographics, clinical outcomes.

interoception2_df <- read.csv("https://raw.githubusercontent.com/RealityBending/InteroceptionScale/refs/heads/main/study2/data/data_participants.csv") |>
  mutate(Source = "InteroceptionScale_Study2") |>
  rename(
    # Deficit metacluster
    MINT_Urin_13 = MINT_Deficit_Urin_1,
    MINT_Urin_14 = MINT_Deficit_Urin_2,
    MINT_Urin_15 = MINT_Deficit_Urin_3,
    MINT_CaCo_10 = MINT_Deficit_CaCo_4,
    MINT_CaCo_11 = MINT_Deficit_CaCo_5,
    MINT_CaCo_12 = MINT_Deficit_CaCo_6,
    MINT_Olfa_22 = MINT_Deficit_Olfa_10,
    MINT_Olfa_23 = MINT_Deficit_Olfa_11,
    MINT_Olfa_24 = MINT_Deficit_Olfa_12,
    MINT_Sati_19 = MINT_Deficit_Sati_13,
    MINT_Sati_20 = MINT_Deficit_Sati_14,
    MINT_Sati_21 = MINT_Deficit_Sati_15,
    # Awareness metacluster
    MINT_SexS_7 = MINT_Awareness_SexS_19,
    MINT_SexS_9 = MINT_Awareness_SexS_20,
    MINT_SexS_8 = MINT_Awareness_SexS_21,
    MINT_RelA_5 = MINT_Awareness_RelA_28,
    MINT_RelA_4 = MINT_Awareness_RelA_29,
    MINT_RelA_6 = MINT_Awareness_RelA_30,
    MINT_ExAc_3 = MINT_Awareness_ExAc_34,
    MINT_ExAc_1 = MINT_Awareness_ExAc_35,
    MINT_ExAc_2 = MINT_Awareness_ExAc_36,
    # Visceroception metacluster
    MINT_Card_28 = MINT_Sensitivity_Card_37,
    MINT_Card_30 = MINT_Sensitivity_Card_38,
    MINT_Card_29 = MINT_Sensitivity_Card_39,
    MINT_Resp_26 = MINT_Sensitivity_Resp_40,
    MINT_Resp_27 = MINT_Sensitivity_Resp_41,
    MINT_Resp_25 = MINT_Sensitivity_Resp_42,
    MINT_Gast_31 = MINT_Sensitivity_Gast_46,
    MINT_Gast_32 = MINT_Sensitivity_Gast_47,
    MINT_Gast_33 = MINT_Sensitivity_Gast_48,
    MINT_Derm_16 = MINT_Sensitivity_Derm_49,
    MINT_Derm_17 = MINT_Sensitivity_Derm_50,
    MINT_Derm_18 = MINT_Sensitivity_Derm_51
  ) |>
  convert_logical_cols() |>
  drop_unused_mint()  # Remove CaNo, SexO, UrSe, StaS


cat("Interoception Study 2 loaded: N =", nrow(interoception2_df), "\n")
```

Interoception Study 2: 118 excluded for failing at least one attention check and 6 based on multivariate distance (using the same procedure as for study 1). 60 excluded due to missing data from a technical issue.

### 1d. FakeArt Dataset

```{r}
# N=359. Mint items already use MINT_Facet_N naming (matching Study 2 after renaming).

fakeart_df <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeArt/refs/heads/main/data/rawdata_participants.csv") |>
  mutate(Source = "FakeArt")

# compute and remove function
compute_and_remove <- function(fakeart_df, name="BodyAwareness", pattern=name, method="mean") {
  items <- select(fakeart_df, starts_with(pattern), -contains("AttentionCheck"))
  fakeart_df <- fakeart_df[!names(fakeart_df) %in% names(items)]
  if(method == "mean") {
    fakeart_df[[name]] <- rowMeans(items, na.rm=TRUE)
  } else {
    fakeart_df[[name]] <- rowSums(items, na.rm=TRUE)
  }
  fakeart_df
}

# PHQ-4
fakeart_df <- compute_and_remove(fakeart_df, name="PHQ4_Depression", method="mean")
fakeart_df <- compute_and_remove(fakeart_df, name="PHQ4_Anxiety", method="mean")

# VVIQ
fakeart_df <- compute_and_remove(fakeart_df, name="VVIQ_Friend", method="mean")
fakeart_df <- compute_and_remove(fakeart_df, name="VVIQ_Sun", method="mean")
fakeart_df <- compute_and_remove(fakeart_df, name="VVIQ_Shop", method="mean")
fakeart_df <- compute_and_remove(fakeart_df, name="VVIQ_Country", method="mean")

fakeart_df$VVIQ_Total = rowMeans(fakeart_df[c("VVIQ_Friend", "VVIQ_Sun", "VVIQ_Shop", "VVIQ_Country")], na.rm = TRUE) |> 
  datawizard::reverse_scale()  # Reverse the scale (e.g., 0–4 -> 4–0) so that higher = more vivid
fakeart_df <- select(fakeart_df, -VVIQ_Friend, -VVIQ_Sun, -VVIQ_Shop, -VVIQ_Country)

# BAIT
fakeart_df$BAIT_Negative <- rowMeans(select(fakeart_df, BAIT_8_TextIssues, BAIT_2_ImagesIssues, BAIT_3_VideosIssues, BAIT_9_Dangerous, BAIT_13_ArtIssues), na.rm=TRUE)
fakeart_df$BAIT_Positive <- rowMeans(select(fakeart_df, BAIT_5_ImitatingReality, BAIT_12_Benefit, BAIT_4_VideosRealistic, BAIT_1_ImagesRealistic, BAIT_14_ArtRealistic, BAIT_6_EnvironmentReal, BAIT_7_TextRealistic), na.rm=TRUE)
fakeart_df$BAIT_ArtRealistic <- rowMeans(select(mutate(fakeart_df, BAIT_13_ArtIssues = abs(BAIT_13_ArtIssues - 6)), BAIT_13_ArtIssues, BAIT_14_ArtRealistic), na.rm=TRUE)
fakeart_df$BAIT_Total <- rowMeans(select(mutate(fakeart_df, BAIT_Negative = abs(BAIT_Negative - 6)), BAIT_Negative, BAIT_Positive), na.rm=TRUE)

## Exclusions 
exclude <- list()

# Attention checks

fakeart_checks <- data.frame(
  Participant = fakeart_df$Participant,
  A_MINT = ifelse(fakeart_df$MINT_AttentionCheck == 0, 0, 1),
  A_BAIT = ifelse(fakeart_df$BAIT_AttentionCheck == 6, 0, 1)
)
fakeart_checks$fail <- fakeart_checks$A_MINT + fakeart_checks$A_BAIT
exclude$checks <- as.character(fakeart_checks$Participant[fakeart_checks$fail > 0])

fakeart_df <- fakeart_df[!fakeart_df$Participant %in% exclude$checks, ]

cat(
  "We excluded", length(exclude$checks), "participants (",
  insight::format_percent(length(exclude$checks) / nrow(fakeart_df)),
  ") who failed at least one attention check.\n"
)
cat("FakeArt loaded: N =", nrow(fakeart_df), "\n")
```

### 1e. FakeChat Dataset

```{r}
fakechat_df <- read.csv("https://raw.githubusercontent.com/EmmaLBenn/ChatExtracts/refs/heads/main/data/FakeChat_data_participants.csv") |>
  mutate(Source = "FakeChat")

cat("FakeChat loaded: N =", nrow(fakechat_df), "\n")
```
 
FakeChat: 58 participants excluded for failing at least one attention check.
 
### 1f. Merge All Datasets

```{r}
# combine all datasets
datasets <- list(
  interoception1_df,
  interoception2_df,
  fakeart_df,
  fakechat_df
)

mega_df <- bind_rows(datasets)

# datasets <- append(datasets, list(new_dataset))   ## use to add more data if necessary

cat("\n=== FINAL MERGED DATASET ===\n")
cat("Total N =", nrow(mega_df), "\n")

cat("Sources:\n")
print(table(mega_df$Source, useNA = "ifany"))

# Identify MINT items
mint_items <- select(mega_df, starts_with("MINT_")) |>
  select(-MINT_AttentionCheck)

cat("\nMINT items in merged dataset:", length(mint_items), "\n")

# Missingness check by dataset
#mega_df |>
#  group_by(Source) |>
#  summarise(
#    n = n(),
#    pct_missing_mint = round(mean(is.na(across(all_of(mint_items)))), 3),
#    .groups = "drop"
#  ) |>
#  print()
```


## Final Sample

::: {.panel-tabset}

The final sample includes `r report::report_participants(mega_df, age = "Age", sex = NA, gender = "Gender")`.

### Age and Gender

```{r p_demographics_age_gender}
p_age <- mega_df |>
  ggplot(aes(x = Age, fill = Gender)) + 
  geom_histogram(data=mega_df, aes(x = Age, fill=Gender), binwidth = 2) +
  geom_vline(xintercept = mean(df$Age), color = "red", linewidth=1.5, linetype="dashed") +
  scale_fill_viridis_d() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(20, max(df$Age), by = 10 )) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Age", y = "Number of Participants", color = NULL, subtitle = "Distribution of participants' age") +
  theme_modern(axis.title.space = 10) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1)),
    axis.title.x = element_blank()
  )

p_age
```

### Education

```{r p_demographics_education}
#| warning: false
#| error: false

p_edu <- mega_df |>
  mutate(Student = ifelse(is.na(Student), FALSE, Student),
         Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate")) |>
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Student)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level", fill = "Is currently a student") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(0.5), angle = 45, hjust =1),
    axis.title.x = element_blank()
  )

p_edu
```

### Ethnicity

```{r p_demographics_ethnicity}
#| warning: false
#| error: false

p_eth <- mega_df |>
  filter(!is.na(Ethnicity)) |> 
  ggplot(aes(x = "", fill = Ethnicity)) +
  geom_bar() +
  coord_polar("y") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank()
  )
    
p_eth
```

### Mental Health


```{r}
mega_df$Disorders_Psychiatric_Mood <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)
mega_df$Disorders_Psychiatric_MoodTreatment <- ifelse(
  mega_df$Disorders_Psychiatric_Mood & !is.na(mega_df$Disorders_PsychiatricTreatment) & str_detect(mega_df$Disorders_PsychiatricTreatment, "Mood|Antidepressant|Anxiolytic|Psychotherapy"), TRUE, FALSE)

mega_df$Disorders_Psychiatric_Anxiety <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "Panic|GAD|Social Phobia|Phobia"), TRUE, FALSE)
mega_df$Disorders_Psychiatric_AnxietyTreatment <- ifelse(
  mega_df$Disorders_Psychiatric_Anxiety & !is.na(mega_df$Disorders_PsychiatricTreatment) & str_detect(mega_df$Disorders_PsychiatricTreatment, "Anxiolytic|Psychotherapy|Mindfulness"), TRUE, FALSE)

mega_df$Disorders_Psychiatric_Eating <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "Eating"), TRUE, FALSE)
mega_df$Disorders_Psychiatric_Addiction <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "Addiction"), TRUE, FALSE)
mega_df$Disorders_Psychiatric_Borderline <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "BPD"), TRUE, FALSE)

mega_df$Disorders_Psychiatric_Autism <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "ASD"), TRUE, FALSE)
mega_df$Disorders_Psychiatric_ADHD <- ifelse(str_detect(mega_df$Disorders_Psychiatric, "ADHD"), TRUE, FALSE)


# df$Disorders_Psychiatric_Other <- ifelse(str_detect(df$Disorders_Psychiatric, "MDD|GAD|Bipolar"), TRUE, FALSE)

select(mega_df, Participant, Gender, 
             Disorders_Psychiatric_Mood, Disorders_Psychiatric_MoodTreatment,
             Disorders_Psychiatric_Anxiety, Disorders_Psychiatric_AnxietyTreatment,
             Disorders_Psychiatric_Eating, Disorders_Psychiatric_Addiction, Disorders_Psychiatric_Borderline,
             Disorders_Psychiatric_Autism, Disorders_Psychiatric_ADHD) |> 
  pivot_longer(cols = starts_with("Disorders_"), names_to = "Disorder", values_to = "Value") |> 
  mutate(Disorder = str_remove_all(Disorder, fixed("Disorders_Psychiatric_")),
         Disorder = str_replace(Disorder, "MoodTreatment", "Mood Disorder (with treatment)"),
         Disorder = str_replace(Disorder, "AnxietyTreatment", "Anxiety (with treatment)"),
         Disorder = str_replace(Disorder, "Mood$", "Mood Disorder")) |> 
  summarize(N = sum(Value) / nrow(mega_df), .by=c("Gender", "Disorder")) |> 
  mutate(N_tot = sum(N), .by="Disorder") |>
  mutate(Disorder = fct_reorder(Disorder, desc(N_tot)),
         Treatment = ifelse(str_detect(Disorder,  "Treatment"), "With treatment", "Without treatment")) |>
  ggplot(aes(x = Disorder, y = N, fill=Gender)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0), labels=scales::percent) +
  labs(title = "Mood Disorders", y = "Number of Participants", subtitle = "Participants per self-declared mood disorder", x="Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```


### Somatic

```{r}
somatic <- data.frame(
  Hypermobility = ifelse(str_detect(mega_df$Disorders_Somatic, "Hypermobility Syndrome"), TRUE, FALSE),
  Fibromyalgia = ifelse(str_detect(mega_df$Disorders_Somatic, "Fibromyalgia"), TRUE, FALSE),
  ChronicFatigue = ifelse(str_detect(mega_df$Disorders_Somatic, "Chronic Fatigue Syndrome"), TRUE, FALSE),
  ChronicPain = ifelse(str_detect(mega_df$Disorders_Somatic, "Chronic Pain Syndrome"), TRUE, FALSE),
  BackPain = ifelse(str_detect(mega_df$Disorders_Somatic, "Back Pain"), TRUE, FALSE),
  MuscleTension = ifelse(str_detect(mega_df$Disorders_Somatic, "Muscle Tension"), TRUE, FALSE),
  SkinRashes = ifelse(str_detect(mega_df$Disorders_Somatic, "Skin Rashes"), TRUE, FALSE),
  Eczema = ifelse(str_detect(mega_df$Disorders_Somatic, "Eczema"), TRUE, FALSE),
  Psoriasis = ifelse(str_detect(mega_df$Disorders_Somatic, "Psoriasis"), TRUE, FALSE),
   # Dermatological = ifelse(str_detect(df$Disorders_Somatic, "Other Dermatological"), TRUE, FALSE),
  # Musculoskeletal = ifelse(str_detect(df$Disorders_Somatic, "Other Musculoskeletal"), TRUE, FALSE),
  Sjogrens = ifelse(str_detect(mega_df$Disorders_Somatic, "Sjogren"), TRUE, FALSE),
  ChestPain = ifelse(str_detect(mega_df$Disorders_Somatic, "Chest Pain"), TRUE, FALSE),
  Arrhythmia = ifelse(str_detect(mega_df$Disorders_Somatic, "Cardiac Arrhythmia"), TRUE, FALSE),
  Hypertension = ifelse(str_detect(mega_df$Disorders_Somatic, "Hypertension"), TRUE, FALSE),
  Hypotension = ifelse(str_detect(mega_df$Disorders_Somatic, "Hypotension"), TRUE, FALSE),
  IBS = ifelse(str_detect(mega_df$Disorders_Somatic, "IBS"), TRUE, FALSE),
  GERD = ifelse(str_detect(mega_df$Disorders_Somatic, "GERD"), TRUE, FALSE),
  Crohn = ifelse(str_detect(mega_df$Disorders_Somatic, "Crohn"), TRUE, FALSE),
  UlcerativeColitis = ifelse(str_detect(mega_df$Disorders_Somatic, "Ulcerative Colitis"), TRUE, FALSE),
  Celiac = ifelse(str_detect(mega_df$Disorders_Somatic, "Celiac Disease"), TRUE, FALSE),
  Gluten = ifelse(str_detect(mega_df$Disorders_Somatic, "Gluten"), TRUE, FALSE),
  Lactose = ifelse(str_detect(mega_df$Disorders_Somatic, "Lactose"), TRUE, FALSE),
  ShortBreath = ifelse(str_detect(mega_df$Disorders_Somatic, "Shortness of Breath"), TRUE, FALSE),
  Asthma = ifelse(str_detect(mega_df$Disorders_Somatic, "Asthma"), TRUE, FALSE),
  COPD = ifelse(str_detect(mega_df$Disorders_Somatic, "COPD"), TRUE, FALSE),
  SleepApnea = ifelse(str_detect(mega_df$Disorders_Somatic, "Sleep Apnea"), TRUE, FALSE),
  Bronchitis = ifelse(str_detect(mega_df$Disorders_Somatic, "Chronic Bronchitis"), TRUE, FALSE),
  Nausea = ifelse(str_detect(mega_df$Disorders_Somatic, "Nausea/Vomiting"), TRUE, FALSE),
  Dizziness = ifelse(str_detect(mega_df$Disorders_Somatic, "Dizziness/Lightheadedness"), TRUE, FALSE),
  Migraine = ifelse(str_detect(mega_df$Disorders_Somatic, "Migraine"), TRUE, FALSE),
  Neuropathy = ifelse(str_detect(mega_df$Disorders_Somatic, "Neuropathy"), TRUE, FALSE),
  Epilepsy = ifelse(str_detect(mega_df$Disorders_Somatic, "Epilepsy"), TRUE, FALSE),
  MS = ifelse(str_detect(mega_df$Disorders_Somatic, "Multiple Sclerosis"), TRUE, FALSE),
  FrequentUrination = ifelse(str_detect(mega_df$Disorders_Somatic, "Frequent Urination"), TRUE, FALSE),
  Endometriosis = ifelse(str_detect(mega_df$Disorders_Somatic, "Endometriosis"), TRUE, FALSE),
  Cystitis = ifelse(str_detect(mega_df$Disorders_Somatic, "Interstitial Cystitis"), TRUE, FALSE),
  PelvicPain = ifelse(str_detect(mega_df$Disorders_Somatic, "Chronic Pelvic Pain Syndrome"), TRUE, FALSE)
)


somatic |> 
  select(where(~n_distinct(.) > 1)) |> 
  correlation::correlation(redundant = TRUE, p_adjust="none") |>
  correlation::cor_sort() |> 
  mutate(label = ifelse(p < .001, "*", ""),
         label = ifelse(r == 1, "", label)) |> 
  ggplot(aes(x=Parameter1, y=Parameter2, fill=r)) +
  geom_tile() +
  geom_text(aes(label=label), size=3) +
  scale_fill_gradient2(low="blue", high="red", mid="white", midpoint=0) +
  labs(title = "Co-occurence Matrix of Somatic Disorders") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle=45, hjust=1),
        axis.title = element_blank())
```

To facilitate mechanistic interpretation and reduce redundancy, self-reported somatic symptoms were grouped into four non-overlapping clusters based on shared physiological pathways and known etiological mechanisms. The **Afferent Sensitivity** cluster included conditions associated with heightened interoceptive awareness and neurogenic excitability, such as migraine, neuropathy, dizziness, nausea, muscle tension, epilepsy, and frequent urination. The **Central Sensitization** cluster comprised syndromes characterized by chronic pain and fatigue, likely reflecting central amplification of sensory signals and HPA-axis dysregulation; this included fibromyalgia, chronic fatigue syndrome, chronic pain, back pain, pelvic pain, endometriosis, irritable bowel syndrome (IBS), and interstitial cystitis. The **Autonomic Dysfunction** cluster captured disorders linked to dysregulation of the autonomic and cardiopulmonary systems, including joint hypermobility, cardiac arrhythmia, chest pain, shortness of breath, hypo-/hypertension, sleep apnea, chronic obstructive pulmonary disease (COPD), and chronic bronchitis. Finally, the **Immune-Inflammatory** cluster encompassed conditions associated with immune dysregulation, barrier dysfunction, and gut-brain axis disturbance, such as eczema, psoriasis, skin rashes, asthma, celiac disease, gluten and lactose sensitivity, inflammatory bowel diseases (Crohn’s disease and ulcerative colitis), gastroesophageal reflux disease (GERD), multiple sclerosis, and Sjögren’s syndrome. Each symptom was assigned to a single cluster to preserve orthogonality and support subsequent analyses linking symptom profiles to interoceptive processes.

```{r}
# df$Disorders_Somatic_Number <- ifelse(df$Disorders_Somatic == "", 0, str_count(df$Disorders_Somatic, ";")+1)

# Cluster A: Afferent Sensitivity (neurogenic excitability, interoceptive hypervigilance)
mega_df$Disorders_Somatic_AfferentSensitivity_N <- rowSums(somatic[, c(
  "Migraine", "Neuropathy", "MuscleTension", "Dizziness", 
  "Nausea", "Epilepsy", "FrequentUrination"
)])
mega_df$Disorders_Somatic_AfferentSensitivity <- mega_df$Disorders_Somatic_AfferentSensitivity_N >= 1

# Cluster B: Central Sensitization (pain and fatigue syndromes)
mega_df$Disorders_Somatic_CentralSensitization_N <- rowSums(somatic[, c(
  "Fibromyalgia", "ChronicPain", "ChronicFatigue", "BackPain", 
  "PelvicPain", "Endometriosis", "IBS", "Cystitis"
)])
mega_df$Disorders_Somatic_CentralSensitization <- mega_df$Disorders_Somatic_CentralSensitization_N >= 1

# Cluster C: Autonomic Dysfunction (dysautonomia and cardiorespiratory instability)
mega_df$Disorders_Somatic_AutonomicDysfunction_N <- rowSums(somatic[, c(
  "Hypermobility", "Arrhythmia", "ChestPain", "ShortBreath", 
  "Hypotension", "Hypertension", "SleepApnea", "COPD", "Bronchitis"
)])
mega_df$Disorders_Somatic_AutonomicDysfunction <- mega_df$Disorders_Somatic_AutonomicDysfunction_N >= 1

# Cluster D: Immune-Inflammatory (gut-brain axis, autoimmune and barrier-related conditions)
mega_df$Disorders_Somatic_ImmuneInflammatory_N <- rowSums(somatic[, c(
  "SkinRashes", "Eczema", "Psoriasis", "Sjogrens", "Asthma", 
  "Celiac", "Gluten", "Lactose", "Crohn", "UlcerativeColitis", 
  "GERD", "MS"
)])
mega_df$Disorders_Somatic_ImmuneInflammatory <- mega_df$Disorders_Somatic_ImmuneInflammatory_N >= 1




p2 <- mega_df |> 
  select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
  select(ends_with("_N"), Gender, Participant) |>
  pivot_longer(starts_with("Disorders_Somatic_")) |> 
  summarize(n = n(), .by=c("name", "value", "Gender")) |> 
  mutate(name = str_remove(name, "Disorders_Somatic_"),
         name = str_remove(name, "_N")) |> 
  ggplot(aes(x = value, y=n, fill=Gender)) +
  geom_bar(stat="identity") +
  scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Number of Issues") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~name)
p2
# 
# p2 <- mega_df |> 
#   select(starts_with("Disorders_Somatic_"), Gender, Participant) |> 
#   select(-ends_with("_N"), Gender, Participant) |>
#   pivot_longer(starts_with("Disorders_Somatic_")) |> 
#   summarize(n = n(), .by=c("name", "value", "Gender")) |> 
#   filter(value == TRUE) |> 
#   ggplot(aes(x = name, y=n, fill=Gender)) +
#   geom_bar(stat="identity") +
#   scale_fill_manual(values = c("Male"= "#64B5F6", "Female"= "#F06292", "Other"="orange", "Missing"="brown")) +
#   scale_y_continuous(expand = c(0, 0)) +
#   labs(title = "Somatic Issues", y = "Number of Participants", subtitle = "Participants per somatic issues", x="Type") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
# p2
```


:::

## 2. Compute Facet & Metacluster Scores

```{r}
# Lower order facet scores (mean of 3 items each)
mega_df <- mega_df |>
  mutate(
    # Awareness metacluster
    MINT_SexS  = rowMeans(across(starts_with("MINT_SexS_")),  na.rm = TRUE),
    MINT_ExAc  = rowMeans(across(starts_with("MINT_ExAc_")),  na.rm = TRUE),
    MINT_RelA  = rowMeans(across(starts_with("MINT_RelA_")),  na.rm = TRUE),
    # Deficit metacluster
    MINT_CaCo  = rowMeans(across(starts_with("MINT_CaCo_")),  na.rm = TRUE),
    MINT_Urin  = rowMeans(across(starts_with("MINT_Urin_")),  na.rm = TRUE),
    MINT_Derm  = rowMeans(across(starts_with("MINT_Derm_")),  na.rm = TRUE),
    MINT_Sati  = rowMeans(across(starts_with("MINT_Sati_")),  na.rm = TRUE),
    MINT_Olfa  = rowMeans(across(starts_with("MINT_Olfa_")),  na.rm = TRUE),
    # Visceroception metacluster
    MINT_Card  = rowMeans(across(starts_with("MINT_Card_")),  na.rm = TRUE),
    MINT_Resp  = rowMeans(across(starts_with("MINT_Resp_")),  na.rm = TRUE),
    MINT_Gast  = rowMeans(across(starts_with("MINT_Gast_")),  na.rm = TRUE)
  )

# Higher order metacluster scores (mean of facets)
mega_df <- mega_df |>
  mutate(
    MINT_Awareness = rowMeans(across(c(MINT_SexS, MINT_ExAc, MINT_RelA)), na.rm = TRUE),
    MINT_Deficit = rowMeans(across(c(MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa)), na.rm = TRUE),
    MINT_Visceroception = rowMeans(across(c(MINT_Card, MINT_Resp, MINT_Gast)), na.rm = TRUE)
  )

mint_lowerorder <- mega_df |> 
  select(MINT_SexS, MINT_ExAc, MINT_RelA, MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa, MINT_Card, MINT_Resp, MINT_Gast)

mint_higherorder <- mega_df |> 
  select(MINT_Awareness, MINT_Deficit, MINT_Visceroception)

cat("Facet scores computed for N =", nrow(mega_df), "participants\n")
cat("Sources included:", paste(unique(mega_df$Source), collapse = ", "), "\n")

# Structure Validation
names(mint_items) <- str_remove(names(mint_items), "MINT_")
```

### 2b. Item Labels

```{r}
labels <- list(
  MINT_ExAc_1 = 
    "I can always accurately feel when I am about to fart",
  MINT_ExAc_2 = 
    "I can always accurately feel when I am about to sneeze",
  MINT_ExAc_3 = 
    "I can always accurately feel when I am about to burp",
  MINT_RelA_4 = 
    "I always feel in my body if I am relaxed",
  MINT_RelA_5 = 
    "I always know when I am relaxed",
  MINT_RelA_6 =
    "My body is always in the same specific state when I am relaxed",
  MINT_SexS_7 =
    "During sex or masturbation, I often feel very strong sensations coming from my genital areas",
  MINT_SexS_8 =
    "My genital organs are very sensitive to pleasant stimulations",
  MINT_SexS_9 =
    "When I am sexually aroused, I often notice specific sensations in my genital area (e.g., tingling, warmth, wetness, stiffness, pulsations)",
  MINT_CaCo_10 =
    "Sometimes my breathing becomes erratic or shallow and I often don't know why",
  MINT_CaCo_11 =
    "I often feel like I can't get enough oxygen by breathing normally",
  MINT_CaCo_12 =
    "Sometimes my heart starts racing and I often don't know why",
  MINT_Urin_13 =
    "I sometimes feel like I need to urinate or defecate but when I go to the bathroom I produce less than I expected",
  MINT_Urin_14 =
    "I often feel the need to urinate even when my bladder is not full",
  MINT_Urin_15 =
    "Sometimes I am not sure whether I need to go to the toilet or not (to urinate or defecate)",
  MINT_Derm_16 = 
    "In general, my skin is very sensitive",
  MINT_Derm_17 = 
    "My skin is susceptible to itchy fabrics and materials",
  MINT_Derm_18 =
    "I can notice even very subtle stimulations to my skin (e.g., very light touches)",
  MINT_Sati_19 =
    "I don't always feel the need to eat until I am really hungry",
  MINT_Sati_20 =
    "Sometimes I don't realise I was hungry until I ate something ",
  MINT_Sati_21 =
    "I don't always feel the need to drink until I am really thirsty",
  MINT_Olfa_22 = 
    "I often check the smell of my armpits",
  MINT_Olfa_23 = 
    "I often check the smell of my own breath",
  MINT_Olfa_24 = 
    "I often check the smell of my farts",
  MINT_Resp_25 = 
    "In general, I am very sensitive to changes in my breathing",
  MINT_Resp_26 = 
    "I can notice even very subtle changes in my breathing",
  MINT_Resp_27 =
    "I am always very aware of how I am breathing, even when I am calm",
  MINT_Card_28 = 
    "In general, I am very sensitive to changes in my heart rate",
  MINT_Card_29 = 
    "I often notice changes in my heart rate",
  MINT_Card_30 =
    "I can notice even very subtle changes in the way my heart beats",
  MINT_Gast_31 =
    "I can notice even very subtle changes in what my stomach is doing",
  MINT_Gast_32 = 
    "In general, I am very sensitive to what my stomach is doing",
  MINT_Gast_33 =
    "I am always very aware of what my stomach is doing, even when I am calm"
)
```

```{r}
cleanlabels <- function(x, qname=FALSE) {
  is_factor <- FALSE
  if(is.factor(x)) {
    is_factor <- TRUE
    vec <- x
    x <- levels(x)
  }
  
  x <- x |> 
    str_replace("MINT_", ifelse(qname, "MINT - ", "")) |> 
    str_replace("_", " - ") 
  
  if(is_factor) {
    levels(vec) <- x
    return(vec)
  }
  x
}
```


## 3. Replicate EGA from InteroceptionScale Study 2

```{r}
#| eval: false

ega <- bootEGA(
  mint_items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  # consensus.method = "lowest_tefi",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, 
  ncores = 4)

save(ega, file="models/ega.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega.RData")

EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega, IS.plot=FALSE)
p_stab <- plot(itemstability)
p_stab

EGAnet::genTEFI(ega$EGA)


ega$EGA
```

Lower-order dimensions are very stable. Higher-order dimensions are mostly stable but a few items show ambiguity (Derm 16, 17 & 18, Olfa 22, 23 & 24).

Deficit metacluster is less consistent (0.546) - less robust, possibly due to heterogeneous items

Network characteristics:
Lower-order network - 33 nodes, 194 edges, edge density 0.367.
Most edges are small (0.071 ± 0.108), range -0.088 to 0.48 → mostly weak to moderate partial correlations.
Higher-order network - 11 nodes (representing lower-order communities), 36 edges, edge density 0.655 → more connected than lower-order network.
Edge weights are slightly stronger on average (0.096 ± 0.103), some negative → stronger inter-community relationships, but some inverse relationships.

Total Entropy Fit Index:
More negative TEFI = better fit of bifactor/hierarchical model.
Results suggest a bifactor structure fits better than just correlated lower-order factors.


## 4. Transpose data to make participants into columns

```{r}
transposed_mint_lowerorder <- as.data.frame(t(mint_lowerorder))
```


## 5. Clustering

```{r}
# Scale data

mint_scaled <- as.data.frame(scale(mint_lowerorder))

```


### 5a. Estimate Number of Clusters

```{r}
cluster_estimate <- n_clusters(
  mint_scaled,
  package = c("all"),
  n_max = 10,
)
cluster_estimate

plot(cluster_estimate)
```

Method Agreement Procedure:
The choice of 2 clusters is supported by 9 (31.03%) methods out of 29 (Elbow, Silhouette, Ch, CCC, DB, Duda, Pseudot2, Beale, Mcclain).

3 clusters also appears to be equally supported.

Cluster estimation suggested 2–3 clusters. While a 2-cluster solution captured a general low vs high distinction, a 3-cluster solution revealed an additional subgroup characterised by globally reduced interoceptive sensitivity. Given its interpretability and theoretical relevance, the 3-cluster solution was retained.

### 5b. Kmeans

```{r}
# 2 clusters

rez_kmeans2 <- cluster_analysis(mint_scaled, n = 2, method = "kmeans")

rez_kmeans2

plot(summary(rez_kmeans2)) # Visualize cluster centers

predict(rez_kmeans2) # Get clusters


# 3 clusters

rez_kmeans3 <- cluster_analysis(mint_scaled, n = 3, method = "kmeans")

rez_kmeans3

plot(summary(rez_kmeans3)) # Visualize cluster centers

predict(rez_kmeans3) # Get clusters
```

Cluster 1 (Low across most facets) - Globally low interoception
Especially low on:
Cardioception (-0.84)
Respiroception (-0.82)
Dermal Hypersensitivity (-0.58)

Cluster 2 - High awareness/relatively low confusion
High on: Awareness facets (SexS, ExAc, RelA) and Visceroception (Card, Resp, Gast)
But slightly low on deficit facets (CaCo, Urin)

Cluster 3 (Mixed pattern) - High confusion/dysregulated interoception
High on Deficit facets (CaCo = +0.94, Urin = +0.71)
Moderate/high elsewhere


### 5c. Hierarchical Clustering

```{r}
rez_hclust <- cluster_analysis(mint_scaled, n = 3, method = "hclust")

rez_hclust

# Visualize
plot(rez_hclust) + theme_modern()

predict(rez_hclust)
```

### 5d. Hierarchical K-Means

```{r}
rez_hkmeans <- cluster_analysis(mint_scaled, n = 3, method = "hkmeans")

rez_hkmeans
```

### 5e. K-Medoids (PAM)

```{r}
rez_pam <- cluster_analysis(mint_scaled, n = 3, method = "pam")

rez_pam

# Visualize
plot(rez_pam) + theme_modern()
```

### 5f. Bootstrapped Hierarchical Clustering

```{r}
# WARNING - SLOW

#rez_hclust2 <- cluster_analysis(mint_scaled,
#  n = NULL,
#  method = "hclust",
#  iterations = 500,
#  ci = 0.90
#)

#rez_hclust2
#plot(rez_hclust2) + theme_modern() # Visualize
```

### 5g. DBSCAN - REMOVE?

```{r}
eps <- n_clusters_dbscan(mint_scaled, min_size = 0.1)
eps
plot(eps) # Find the 'elbow' of the curve: 4.25
```

```{r}
rez_dbscan <- cluster_analysis(mint_scaled, method = "dbscan", dbscan_eps = 4.25)

rez_dbscan
plot(rez_dbscan) + theme_modern() # Visualize
```

### 5h. Hierarchical DBSCAN - REMOVE?

```{r}
rez_hdbscan <- cluster_analysis(mint_scaled, method = "hdbscan")

rez_hdbscan

# Visualize
plot(rez_hdbscan) + theme_modern()
```

### 5i. K-Medoids with estimation of number of clusters (pamk)

```{r}
rez_pamk <- cluster_analysis(mint_scaled, method = "pamk")

rez_pamk

# Visualize
plot(rez_pamk) + theme_modern()
```

### 5j. Mixture - REMOVE?

```{r}
rez_mixture <- cluster_analysis(mint_scaled, method = "mixture")

rez_mixture

# Visualize
plot(rez_mixture) + theme_modern()
```


## 6. Metaclustering

```{r}
list_of_results <- list(
  rez_kmeans3, rez_hclust, rez_hkmeans, rez_pam, rez_pamk
)

probability_matrix <- cluster_meta(list_of_results)

# Plot the matrix as a reordered heatmap
heatmap(unclass(probability_matrix),
  scale = "none",
  col = grDevices::hcl.colors(256, palette = "inferno")
)
```

