---
title: "InteroceptionProfiles - Data Analysis (Emma)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    toc-depth: 3
execute: 
  cache: true
---

# 1. Load packages and data

```{r}
#| message: false
#| warning: false

library(sjmisc)
library(patchwork)
library(ggside)
library(effectsize)
library(EGAnet)
library(nnet)
library(DescTools)
library(broom)
library(mclust)
library(officer)
library(pheatmap)
library(factoextra)
library(tidygraph)
library(ggraph)
library(scales)
library(car)
library(fmsb)
library(ggpubr)
library(ggdist)
library(ggtext)
library(flextable)
library(tidyverse)
library(easystats)

set.seed(42)

mega_df <- read.csv("https://raw.githubusercontent.com/EmmaLBenn/ChatExtracts/refs/heads/main/data/Emma_dissertationdata.csv")
```


# 2. Compute Facet & Metacluster Scores

```{r}
mint_items <- select(mega_df, starts_with("MINT_")) |>
  select(-MINT_AttentionCheck)

# Lower order facet scores (mean of 3 items each)
mega_df <- mega_df |>
  mutate(
    # Awareness metacluster
    MINT_SexS  = rowMeans(across(starts_with("MINT_SexS_")),  na.rm = TRUE),
    MINT_ExAc  = rowMeans(across(starts_with("MINT_ExAc_")),  na.rm = TRUE),
    MINT_RelA  = rowMeans(across(starts_with("MINT_RelA_")),  na.rm = TRUE),
    # Deficit metacluster
    MINT_CaCo  = rowMeans(across(starts_with("MINT_CaCo_")),  na.rm = TRUE),
    MINT_UrIn  = rowMeans(across(starts_with("MINT_UrIn_")),  na.rm = TRUE),
    MINT_Derm  = rowMeans(across(starts_with("MINT_Derm_")),  na.rm = TRUE),
    MINT_Sati  = rowMeans(across(starts_with("MINT_Sati_")),  na.rm = TRUE),
    MINT_Olfa  = rowMeans(across(starts_with("MINT_Olfa_")),  na.rm = TRUE),
    # Visceroception metacluster
    MINT_Card  = rowMeans(across(starts_with("MINT_Card_")),  na.rm = TRUE),
    MINT_Resp  = rowMeans(across(starts_with("MINT_Resp_")),  na.rm = TRUE),
    MINT_Gast  = rowMeans(across(starts_with("MINT_Gast_")),  na.rm = TRUE)
  )

# Higher order metacluster scores (mean of facets)
mega_df <- mega_df |>
  mutate(
    MINT_Awareness = rowMeans(across(c(MINT_SexS, MINT_ExAc, MINT_RelA)), na.rm = TRUE),
    MINT_Deficit = rowMeans(across(c(MINT_CaCo, MINT_UrIn, MINT_Derm, MINT_Sati, MINT_Olfa)), na.rm = TRUE),
    MINT_Visceroception = rowMeans(across(c(MINT_Card, MINT_Resp, MINT_Gast)), na.rm = TRUE)
  )

mint_lowerorder <- mega_df |> 
  select(MINT_SexS, MINT_ExAc, MINT_RelA, MINT_CaCo, MINT_UrIn, MINT_Derm, MINT_Sati, MINT_Olfa, MINT_Card, MINT_Resp, MINT_Gast)

mint_higherorder <- mega_df |> 
  select(MINT_Awareness, MINT_Deficit, MINT_Visceroception)

cat("Facet scores computed for N =", nrow(mega_df), "participants\n")
cat("Sources included:", paste(unique(mega_df$Source), collapse = ", "), "\n")

# Structure Validation
names(mint_items) <- str_remove(names(mint_items), "MINT_")
```


# 3. Replicate EGA from InteroceptionScale Study 2

```{r}
#| eval: false

ega <- bootEGA(
  mint_items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  # consensus.method = "lowest_tefi",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3,
  ncores = 4)

save(ega, file="models/ega.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega.RData")

EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega, IS.plot = FALSE)
p_stab <- plot(itemstability)

p_stab <- p_stab +
  patchwork::plot_annotation(
    title = "Exploratory Graphical Analysis (EGA)",
    subtitle = "Bootstrapped replication of clusters (Method = Leiden)",
    theme = theme(
      plot.title = element_text(face = "bold", hjust = 0, size = rel(1.4)),
      plot.subtitle = element_text(hjust = 0, size = rel(1))
    )
  )

ggsave("figures/ega_stability.png", p_stab, width = 14, height = 12, dpi = 300)

EGAnet::genTEFI(ega$EGA)
ega$EGA
```


# 4. Clustering

```{r}
# Scale data
mint_scaled <- as.data.frame(scale(mint_lowerorder))
```

## 4.1. Estimate Number of Clusters

```{r}
cluster_estimate <- n_clusters(
  mint_scaled,
  package = c("all"),
  n_max = 10,
)
cluster_estimate

plot(cluster_estimate)
```

Method Agreement Procedure:
The choice of 3 clusters is supported by 10 (34.48%) methods out of 29 (Hartigan, Scott, trcovw, Tracew, Friedman, Rubin, Ratkowsky, Ball, SDindex, Mixture (VVE)).

```{r}
# Elbow plot
p_elbow <- fviz_nbclust(mint_scaled, FUNcluster = cluster::pam, method = "wss", k.max = 8) +
  labs(title = "Elbow Method — Optimal Number of Clusters") +
  theme_classic()

# Silhouette plot
p_sil <- fviz_nbclust(mint_scaled, FUNcluster = cluster::pam, method = "silhouette", k.max = 8) +
  labs(title = "Silhouette Method — Optimal Number of Clusters") +
  theme_classic()

p_elbow + p_sil

ggsave("figures/fig_elbow.png", plot = p_elbow, width = 5, height = 6, dpi = 300)
ggsave("figures/fig_silhouette.png", plot = p_sil, width = 5, height = 6, dpi = 300)
```

## 4.2. Kmeans

```{r}
# 3 clusters

rez_kmeans3 <- cluster_analysis(mint_scaled, n = 3, method = "kmeans")

rez_kmeans3

plot(rez_kmeans3)

predict(rez_kmeans3)
```


## 4.3. Hierarchical Clustering

```{r}
rez_hclust <- cluster_analysis(mint_scaled, n = 3, method = "hclust")

rez_hclust

plot(rez_hclust) + theme_modern()

predict(rez_hclust)
```

## 4.4. Hierarchical K-Means

```{r}
rez_hkmeans <- cluster_analysis(mint_scaled, n = 3, method = "hkmeans")

rez_hkmeans

plot(rez_hkmeans) + theme_modern()
```

## 4.5. K-Medoids (PAM)

```{r}
rez_pam <- cluster_analysis(mint_scaled, n = 3, method = "pam")

rez_pam

plot(rez_pam) + theme_modern()
```

## 4.6. DBSCAN

```{r}
# eps <- n_clusters_dbscan(mint_scaled, min_size = 0.1)
# eps
# 
# plot(eps) # Find the 'elbow' of the curve: 4.25
```

```{r}
# rez_dbscan <- cluster_analysis(mint_scaled, method = "dbscan", dbscan_eps = 4.25)
# 
# rez_dbscan
# 
# plot(rez_dbscan) + theme_modern()
```

Suggests the existence of 1 cluster.
The 2 clusters accounted for 1.47% of the total variance of the original data.

Removed due to not fitting well with the data.

## 4.7. K-Medoids with estimation of number of clusters (pamk)

```{r}
rez_pamk <- cluster_analysis(mint_scaled, method = "pamk")

rez_pamk

plot(rez_pamk) + theme_modern()
```

The 2 clusters accounted for 17.50% of the total variance of the original data.

## 4.8. Mixture

```{r}
rez_mixture <- cluster_analysis(mint_scaled, method = "mixture")

rez_mixture

plot(rez_mixture) + theme_modern()
```

The 4 clusters accounted for 13.43% of the total variance of the original data.


# 5. Metaclustering

```{r}
list_of_results <- list(
  rez_kmeans3, rez_hclust, rez_hkmeans, rez_pam, rez_pamk, rez_mixture
 )

probability_matrix <- cluster_meta(list_of_results)

# Plot the matrix as a reordered heatmap
heatmap <- stats::heatmap(probability_matrix,
                          scale = "none",
                          col = grDevices::hcl.colors(256, palette = "inferno")
                          )
heatmap
```

# 6. Extract participants from clusters

K-Medoids (PAM) clustering was chosen as the final clustering solution due to its similarity to the meta-clustering solution and its cluster stability.

```{r}
cluster_assignment <- predict(rez_pam)

mega_df <- mega_df |>
  mutate(
    Clusters = factor(
      cluster_assignment,
      levels = c(1, 2, 3),
      labels = c("Cluster 1", "Cluster 2", "Cluster 3")
    )
  )
```

## 6.1. Heatmap with K-Medoids Overlay

```{r}
order_meta <- heatmap$rowInd

clusters_ord <- mega_df$Clusters[order_meta]

cluster_cols <- c(
  "Cluster 1" = "#6495ed",
  "Cluster 2" = "#64ca84",
  "Cluster 3" = "#b96469"
)

side_cols <- cluster_cols[clusters_ord]

png("figures/overlay_heatmap.png", width = 2800, height = 2000, res = 300)

par(mar = c(2, 2, 10, 2),
oma = c(0, 0, 0, 8))

overlay_heatmap <- heatmap(
  probability_matrix[order_meta, order_meta],
  scale = "none",
  col = grDevices::hcl.colors(256, "inferno"),
  ColSideColors = side_cols,
  labRow = NA,
  labCol = NA,
  main = ""
)

cluster_numeric <- as.numeric(clusters_ord)
boundaries <- cumsum(table(cluster_numeric))
abline(h = nrow(probability_matrix) - boundaries, col = "white", lwd = 2)
abline(v = boundaries, col = "white", lwd = 2)

legend(
  x = "topright",
  inset = c(-0.25, 0.05),
  legend = c(
    "Interoceptive Baseline (Cluster 1)",
    "Visceroceptive Awareness (Cluster 2)",
    "Dysregulated Interoception (Cluster 3)"
  ),
  fill = cluster_cols,
  border = NA,
  bty = "n",
  cex = 0.7,
  xpd = NA
)

dev.off()
```


# 7. Analysis

## 7.1. Age
```{r}
# Effect size: small = .01, medium = .06, large = .14 

aov_age <- aov(Age ~ Clusters, data = mega_df)
summary(aov_age)
eta_squared(aov_age)  # effect size

# Post-hoc test
tukey_age <- TukeyHSD(aov_age)
print(tukey_age)

```


## 7.2. Depression & Anxiety (PHQ-4)

```{r}
# Depression

aov_dep_phq <- aov(PHQ4_Depression ~ Clusters, data = mega_df)
summary(aov_dep_phq)
eta_squared(aov_dep_phq)  # effect size

# Post-hoc test
tukey_dep <- TukeyHSD(aov_dep_phq)
print(tukey_dep)


# Anxiety

aov_anx_phq <- aov(PHQ4_Anxiety ~ Clusters, data = mega_df)
summary(aov_anx_phq)
eta_squared(aov_anx_phq)  # effect size

# Post-hoc test
tukey_anx <- TukeyHSD(aov_anx_phq)
print(tukey_anx)

```


## 7.3. Gender

```{r}
# Cramér's V (2x3 table, df =2): small = .07, medium = .21, large = .35

gender_df <- mega_df |> filter(Gender != "Other")

# Chi-square
chi_gend <- chisq.test(gender_df$Gender, gender_df$Clusters)
chi_gend

# Effect size (Cramér's V)
cramers_v(table(gender_df$Gender, gender_df$Clusters))

# Observed frequencies
chi_gend$observed

# Expected frequencies - check <5
chi_gend$expected


# Row percentages (% of each gender in each cluster)
prop.table(table(gender_df$Gender, gender_df$Clusters), margin = 1) * 100

# Column percentages (% of each cluster that is each gender)
prop.table(table(gender_df$Gender, gender_df$Clusters), margin = 2) * 100
```


## 7.4. Mental Health Conditions (Diagnosed)

### 7.4.1. Mood

```{r}
# Remove NAs
mood_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Mood))

# Chi-square
chi_mood <- chisq.test(mood_df$Disorders_Psychiatric_Mood, mood_df$Clusters)
chi_mood

# Effect size (Cramér's V)
cramers_v(table(mood_df$Disorders_Psychiatric_Mood, mood_df$Clusters))

# Observed frequencies
chi_mood$observed

# Expected frequencies - check <5
chi_mood$expected
```


### 7.4.2. Anxiety

```{r}
# Remove NAs
anxiety_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Anxiety))

# Chi-square
chi_anxiety <- chisq.test(anxiety_df$Disorders_Psychiatric_Anxiety, anxiety_df$Clusters)
chi_anxiety

# Effect size (Cramér's V)
cramers_v(table(anxiety_df$Disorders_Psychiatric_Anxiety, anxiety_df$Clusters))

# Observed frequencies
chi_anxiety$observed

# Expected frequencies - check <5
chi_anxiety$expected
```


### 7.4.3. Eating Disorders

```{r}
# Remove NAs
eating_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Eating))

# Chi-square
chi_eating <- chisq.test(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters)
chi_eating

# Effect size (Cramér's V)
cramers_v(table(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters))

# Observed frequencies
chi_eating$observed

# Expected frequencies - check <5
chi_eating$expected
```


```{r}
fisher.test(table(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters), 
            simulate.p.value = TRUE)
```


### 7.4.4 Autism

```{r}
# Remove NAs
autism_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Autism))

# Chi-square
chi_autism <- chisq.test(autism_df$Disorders_Psychiatric_Autism, autism_df$Clusters)
chi_autism

# Effect size (Cramér's V)
cramers_v(table(autism_df$Disorders_Psychiatric_Autism, autism_df$Clusters))

# Observed frequencies
chi_autism$observed

# Expected frequencies - check <5
chi_autism$expected
```


### 7.4.5. ADHD

```{r}
# Remove NAs
adhd_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_ADHD))

# Chi-square
chi_adhd <- chisq.test(adhd_df$Disorders_Psychiatric_ADHD, adhd_df$Clusters)
chi_adhd

# Effect size (Cramér's V)
cramers_v(table(adhd_df$Disorders_Psychiatric_ADHD, adhd_df$Clusters))

# Observed frequencies
chi_adhd$observed

# Expected frequencies - check for <5
chi_adhd$expected
```


## 7.5. Somatic Health Conditions (Diagnosed)

### 7.5.1. Afferent Sensitivity

```{r}
# Remove NAs
as_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_AfferentSensitivity))

# Chi-square
chi_as <- chisq.test(as_df$Disorders_Somatic_AfferentSensitivity, as_df$Clusters)
chi_as

# Effect size (Cramér's V)
cramers_v(table(as_df$Disorders_Somatic_AfferentSensitivity, as_df$Clusters))

# Observed frequencies
chi_as$observed

# Expected frequencies - check for <5
chi_as$expected
```


### 7.5.2. Central Sensitisation

```{r}
# Remove NAs
cs_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_CentralSensitization))

# Chi-square
chi_cs <- chisq.test(cs_df$Disorders_Somatic_CentralSensitization, cs_df$Clusters)
chi_cs

# Effect size (Cramér's V)
cramers_v(table(cs_df$Disorders_Somatic_CentralSensitization, cs_df$Clusters))

# Observed frequencies
chi_cs$observed

# Expected frequencies - check for <5
chi_cs$expected
```


### 7.5.3. Autonomic Dysfunction

```{r}
# Remove NAs
ad_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_AutonomicDysfunction))

# Chi-square
chi_ad <- chisq.test(ad_df$Disorders_Somatic_AutonomicDysfunction, ad_df$Clusters)
chi_ad

# Effect size (Cramér's V)
cramers_v(table(ad_df$Disorders_Somatic_AutonomicDysfunction, ad_df$Clusters))

# Observed frequencies
chi_ad$observed

# Expected frequencies - check for <5
chi_ad$expected
```


### 7.5.4. Immune-Inflammatory

```{r}
# Remove NAs
ii_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_ImmuneInflammatory))

# Chi-square
chi_ii <- chisq.test(ii_df$Disorders_Somatic_ImmuneInflammatory, ii_df$Clusters)
chi_ii

# Effect size (Cramér's V)
cramers_v(table(ii_df$Disorders_Somatic_ImmuneInflammatory, ii_df$Clusters))

# Observed frequencies
chi_ii$observed

# Expected frequencies - check for <5
chi_ii$expected
```


# 8. Multinomial Logistic Regression

```{r}
gender_df$Male <- rec(gender_df$Gender, rec = "Male = 1; Female = 0")

gender_df$Clusters_relev <- relevel(gender_df$Clusters, ref = "Cluster 1")

multinom_model <- multinom(
  Clusters_relev ~ Age + PHQ4_Anxiety + Gender +
    Disorders_Psychiatric_Mood +
    Disorders_Somatic_AfferentSensitivity +
    Disorders_Somatic_ImmuneInflammatory,
  data = gender_df,
  model = TRUE
)

model_data <- multinom_model$model

# P-values from Wald z-scores
z <- summary(multinom_model)$coefficients / summary(multinom_model)$standard.errors
p_values <- (1 - pnorm(abs(z), 0, 1)) * 2
round(p_values, 3)

# Odds ratios and CIs
tidy(multinom_model, exponentiate = TRUE, conf.int = TRUE) |>
  filter(term != "(Intercept)") |>
  mutate(
    OR_CI = paste0(round(estimate, 2), " [", round(conf.low, 2), ", ", round(conf.high, 2), "]"),
    p_fmt = ifelse(p.value < .001, "< .001", paste0("= ", round(p.value, 3)))
  ) |>
  select(y.level, term, OR_CI, p_fmt)

# Likelihood ratio test
null_model <- multinom(Clusters_relev ~ 1, data = model_data)
lrt_chi <- null_model$deviance - multinom_model$deviance
lrt_df  <- multinom_model$edf - null_model$edf
lrt_p   <- pchisq(lrt_chi, df = lrt_df, lower.tail = FALSE)
cat("LRT: Chi-square =", round(lrt_chi, 2), ", df =", lrt_df, ", p =", format.pval(lrt_p, eps = .001, digits = 3), "\n")

# Pseudo-R²
PseudoR2(multinom_model, which = c("CoxSnell", "Nagelkerke", "McFadden"))

# Mean predicted probabilities by actual cluster
pred_probs <- fitted(multinom_model)

model_data |>
  mutate(P_C1 = pred_probs[, 1], 
         P_C2 = pred_probs[, 2], 
         P_C3 = pred_probs[, 3]) |>
  group_by(Clusters_relev) |>
  summarise(across(c(P_C1, P_C2, P_C3), ~ round(mean(., na.rm = TRUE), 3)))

# Check multicollinearity
lm_check <- lm(
  as.numeric(Clusters_relev) ~
    Age + 
    PHQ4_Anxiety + 
    Gender +
    Disorders_Psychiatric_Mood +
    Disorders_Somatic_AfferentSensitivity +
    Disorders_Somatic_ImmuneInflammatory,
  data = model_data
  )

vif(lm_check)
```


# 9. Figures

## 9.1. Radar chart for cluster profiles

```{r}
facet_order <- c(
  "SexS", "ExAc", "RelA",
  "CaCo", "UrIn", "Derm", "Sati", "Olfa",
  "Card", "Resp", "Gast"
)

radar_wide <- mint_lowerorder |>
  mutate(Cluster = mega_df$Clusters) |>
  group_by(Cluster) |>
  summarise(across(everything(), mean, na.rm = TRUE)) |>
  pivot_longer(-Cluster, names_to = "Facet", values_to = "Value") |>
  mutate(Facet = gsub("^MINT_", "", Facet)) |>
  filter(Facet %in% facet_order) |>
  pivot_wider(names_from = Facet, values_from = Value) |>
  arrange(Cluster) |>
  select(Cluster, all_of(facet_order))

plot_data <- bind_rows(
  setNames(as.data.frame(matrix(rep(6, length(facet_order)), nrow = 1)), facet_order),
  setNames(as.data.frame(matrix(rep(0, length(facet_order)), nrow = 1)), facet_order),
  radar_wide |> select(-Cluster)
)

cluster_labels <- c(
  "Cluster 1 – Interoceptive Baseline",
  "Cluster 2 – Visceroceptive Awareness",
  "Cluster 3 – Dysregulated Interoception"
)

cluster_cols <- c("#6495ed", "#64ca84", "#b96469")

png("figures/radar_clusters.png", width = 3400, height = 1400, res = 300, bg = "white")

par(mfrow = c(1, 3), mar = c(1, 1, 3, 1), oma = c(1, 0, 3, 0))

for (i in 1:3) {
  radarchart(
    plot_data[c(1, 2, i + 2), ],
    axistype    = 1,
    seg         = 3,
    caxislabels = c("0", "2", "4", "6"),
    pcol        = cluster_cols[i],
    pfcol       = adjustcolor(cluster_cols[i], alpha.f = 0.15),
    plwd        = 2.5,
    cglcol      = "grey80",
    cglty       = 1,
    cglwd       = 0.6,
    vlcex       = 0.85,
    axislabcol  = "grey50",
    title       = cluster_labels[i]
  )
}

mtext("MINT Facet Profiles by Cluster", outer = TRUE, side = 3, cex = 1.4, font = 2, line = 0.5)

dev.off()
```

## 9.2. Raincloud plots

```{r}
cluster_cols <- c(
  "Cluster 1" = "#6495ed",
  "Cluster 2" = "#64ca84",
  "Cluster 3" = "#b96469"
)

p_dep <- ggplot(mega_df |>
                  filter(!is.na(PHQ4_Depression)),
                aes(x = Clusters, y = PHQ4_Depression, fill = Clusters)) +
  stat_slab(adjust = 0.5, justification = -0.2) +
  geom_boxplot(width = 0.12, outlier.color = NA, alpha = 0.5) +
  scale_fill_manual(values = cluster_cols) +
  labs(title = "A", y = "Depression (PHQ-4)", x = NULL) +
  coord_flip() +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")

p_anx <- ggplot(mega_df |> 
                  filter(!is.na(PHQ4_Anxiety)),
                aes(x = Clusters, y = PHQ4_Anxiety, fill = Clusters)) +
  stat_slab(adjust = 0.5, justification = -0.2) +
  geom_boxplot(width = 0.12, outlier.color = NA, alpha = 0.5) +
  scale_fill_manual(values = cluster_cols) +
  labs(title = "B", y = "Anxiety (PHQ-4)", x = NULL) +
  coord_flip() +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")

p_age <- ggplot(mega_df,
                aes(x = Clusters, y = Age, fill = Clusters)) +
  stat_slab(adjust = 0.5, justification = -0.2) +
  geom_boxplot(width = 0.12, outlier.color = NA, alpha = 0.5) +
  scale_fill_manual(values = cluster_cols) +
  labs(title = "C", y = "Age (years)", x = NULL) +
  coord_flip() +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")

combined <- p_dep / p_anx / p_age +
  plot_annotation(title = "Continuous Variables by Cluster",
                  theme = theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 14)))

ggsave("figures/rainclouds.png", combined, width = 8, height = 10, dpi = 300)
```

## 9.3. Heatmap for diagnoses

```{r}
diag_data <- mega_df |>
  group_by(Clusters) |>
  summarise(
    "Mood Disorder"         = mean(Disorders_Psychiatric_Mood, na.rm = TRUE),
    "Anxiety Disorder"      = mean(Disorders_Psychiatric_Anxiety, na.rm = TRUE),
    "Eating Disorder"       = mean(Disorders_Psychiatric_Eating, na.rm = TRUE),
    "ASD"                   = mean(Disorders_Psychiatric_Autism, na.rm = TRUE),
    "ADHD"                  = mean(Disorders_Psychiatric_ADHD, na.rm = TRUE),
    "Afferent Sensitivity"  = mean(Disorders_Somatic_AfferentSensitivity, na.rm = TRUE),
    "Central Sensitisation" = mean(Disorders_Somatic_CentralSensitization, na.rm = TRUE),
    "Autonomic Dysfunction" = mean(Disorders_Somatic_AutonomicDysfunction, na.rm = TRUE),
    "Immune-Inflammatory"   = mean(Disorders_Somatic_ImmuneInflammatory, na.rm = TRUE)
  ) |>
  pivot_longer(-Clusters, names_to = "Condition", values_to = "Proportion") |>
  mutate(
    Proportion = Proportion * 100,
    Category = case_when(
      Condition %in% c("Mood Disorder", "Anxiety Disorder", "Eating Disorder", "ASD", "ADHD") ~ "Psychiatric",
      TRUE ~ "Somatic"
    ),
    Condition = factor(Condition, levels = c(
      "Mood Disorder", "Anxiety Disorder", "Eating Disorder", "ASD", "ADHD", "Afferent Sensitivity", "Central Sensitisation", "Autonomic Dysfunction", "Immune-Inflammatory"
    )),
    Clusters = factor(Clusters,
      levels = c("Cluster 1", "Cluster 2", "Cluster 3"),
      labels = c("Cluster 1\nInteroceptive Baseline",
                 "Cluster 2\nVisceroceptive Awareness",
                 "Cluster 3\nDysregulated Interoception"))
  )

ggplot(diag_data, aes(x = Clusters, y = Condition, fill = Proportion)) +
  geom_tile(colour = "white", linewidth = 0.8) +
  geom_text(aes(label = sprintf("%.0f%%", Proportion)),
            size = 3.2, colour = "white", fontface = "bold") +
  facet_grid(Category ~ ., scales = "free_y", space = "free_y") +
  scale_fill_gradient(low = "#f7f0f0", high = "#330066",
                      name = "Prevalence (%)") +
  labs(title = "Prevalence of Conditions by Cluster",
       x = NULL, y = NULL) +
  theme_classic(base_size = 12) +
  theme(
    strip.background = element_blank(),
    strip.text.y     = element_text(face = "bold", angle = 0, hjust = 0),
    legend.position  = "bottom",
    axis.text.x      = element_text(face = "bold"),
    panel.spacing    = unit(0.4, "lines"),
    plot.title       = element_text(face = "bold", hjust = 0.5)
  )

ggsave("figures/diagnoses_heatmap.png", width = 8, height = 7, dpi = 300)
```

## 9.3. Clustering Methods Comparison

```{r}
method_cols <- c("#4a2377", "#f55f74", "#0d7d87", "#8cc5e3")

p_km <- plot(rez_kmeans3) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "k-Means (k = 3)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

p_hc <- plot(rez_hclust) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "Hierarchical Clustering\n— Ward's Method (k = 3)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

p_hkm <- plot(rez_hkmeans) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "Hierarchical k-Means (k = 3)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

p_pam <- plot(rez_pam) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "k-Medoids PAM (k = 3)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

p_pamk <- plot(rez_pamk) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "K-Medoids with estimation of\nnumber of clusters - pamk (k = 2)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

p_mix <- plot(rez_mixture) + 
  theme_classic(base_size = 10) +
  scale_colour_manual(values = method_cols) +
  scale_fill_manual(values = method_cols) +
  labs(title = "Gaussian Mixture Modelling (k = 4)") +
  theme(plot.title = element_text(face = "bold", hjust = 0.5),
        legend.position = "none")

combined <- (p_km | p_hc | p_hkm) / (p_pam | p_pamk | p_mix) +
  plot_annotation(
    title = "Clustering Solutions Across Methods",
    theme = theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 13))
  )

ggsave("figures/cluster_comparison.png", combined, width = 14, height = 9, dpi = 300)
```

