---
title: "InteroceptionProfiles - Data Analysis (Emma)"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: false
    toc: true
    toc-depth: 3
execute: 
  cache: true
---

# 1. Load packages and data

```{r}
#| message: false
#| warning: false

library(sjmisc)
library(cluster)
library(factoextra)
library(mclust)
library(pvclust)
library(NbClust)
library(dbscan)
library(kernlab)
library(apcluster)
library(nnet)
library(umap)
library(Rtsne)
library(EGAnet)
library(ggdist)
library(effectsize)
library(tidyverse)
library(easystats)

set.seed(42)

mega_df <- read.csv("https://raw.githubusercontent.com/EmmaLBenn/ChatExtracts/refs/heads/main/data/Emma_dissertationdata.csv")
```


# 2. Compute Facet & Metacluster Scores

```{r}
mint_items <- select(mega_df, starts_with("MINT_")) |>
  select(-MINT_AttentionCheck)

# Lower order facet scores (mean of 3 items each)
mega_df <- mega_df |>
  mutate(
    # Awareness metacluster
    MINT_SexS  = rowMeans(across(starts_with("MINT_SexS_")),  na.rm = TRUE),
    MINT_ExAc  = rowMeans(across(starts_with("MINT_ExAc_")),  na.rm = TRUE),
    MINT_RelA  = rowMeans(across(starts_with("MINT_RelA_")),  na.rm = TRUE),
    # Deficit metacluster
    MINT_CaCo  = rowMeans(across(starts_with("MINT_CaCo_")),  na.rm = TRUE),
    MINT_Urin  = rowMeans(across(starts_with("MINT_Urin_")),  na.rm = TRUE),
    MINT_Derm  = rowMeans(across(starts_with("MINT_Derm_")),  na.rm = TRUE),
    MINT_Sati  = rowMeans(across(starts_with("MINT_Sati_")),  na.rm = TRUE),
    MINT_Olfa  = rowMeans(across(starts_with("MINT_Olfa_")),  na.rm = TRUE),
    # Visceroception metacluster
    MINT_Card  = rowMeans(across(starts_with("MINT_Card_")),  na.rm = TRUE),
    MINT_Resp  = rowMeans(across(starts_with("MINT_Resp_")),  na.rm = TRUE),
    MINT_Gast  = rowMeans(across(starts_with("MINT_Gast_")),  na.rm = TRUE)
  )

# Higher order metacluster scores (mean of facets)
mega_df <- mega_df |>
  mutate(
    MINT_Awareness = rowMeans(across(c(MINT_SexS, MINT_ExAc, MINT_RelA)), na.rm = TRUE),
    MINT_Deficit = rowMeans(across(c(MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa)), na.rm = TRUE),
    MINT_Visceroception = rowMeans(across(c(MINT_Card, MINT_Resp, MINT_Gast)), na.rm = TRUE)
  )

mint_lowerorder <- mega_df |> 
  select(MINT_SexS, MINT_ExAc, MINT_RelA, MINT_CaCo, MINT_Urin, MINT_Derm, MINT_Sati, MINT_Olfa, MINT_Card, MINT_Resp, MINT_Gast)

mint_higherorder <- mega_df |> 
  select(MINT_Awareness, MINT_Deficit, MINT_Visceroception)

cat("Facet scores computed for N =", nrow(mega_df), "participants\n")
cat("Sources included:", paste(unique(mega_df$Source), collapse = ", "), "\n")

# Structure Validation
names(mint_items) <- str_remove(names(mint_items), "MINT_")
```

## 2.1. Item Labels

```{r}
labels <- list(
  MINT_ExAc_1 = 
    "I can always accurately feel when I am about to fart",
  MINT_ExAc_2 = 
    "I can always accurately feel when I am about to sneeze",
  MINT_ExAc_3 = 
    "I can always accurately feel when I am about to burp",
  MINT_RelA_4 = 
    "I always feel in my body if I am relaxed",
  MINT_RelA_5 = 
    "I always know when I am relaxed",
  MINT_RelA_6 =
    "My body is always in the same specific state when I am relaxed",
  MINT_SexS_7 =
    "During sex or masturbation, I often feel very strong sensations coming from my genital areas",
  MINT_SexS_8 =
    "My genital organs are very sensitive to pleasant stimulations",
  MINT_SexS_9 =
    "When I am sexually aroused, I often notice specific sensations in my genital area (e.g., tingling, warmth, wetness, stiffness, pulsations)",
  MINT_CaCo_10 =
    "Sometimes my breathing becomes erratic or shallow and I often don't know why",
  MINT_CaCo_11 =
    "I often feel like I can't get enough oxygen by breathing normally",
  MINT_CaCo_12 =
    "Sometimes my heart starts racing and I often don't know why",
  MINT_Urin_13 =
    "I sometimes feel like I need to urinate or defecate but when I go to the bathroom I produce less than I expected",
  MINT_Urin_14 =
    "I often feel the need to urinate even when my bladder is not full",
  MINT_Urin_15 =
    "Sometimes I am not sure whether I need to go to the toilet or not (to urinate or defecate)",
  MINT_Derm_16 = 
    "In general, my skin is very sensitive",
  MINT_Derm_17 = 
    "My skin is susceptible to itchy fabrics and materials",
  MINT_Derm_18 =
    "I can notice even very subtle stimulations to my skin (e.g., very light touches)",
  MINT_Sati_19 =
    "I don't always feel the need to eat until I am really hungry",
  MINT_Sati_20 =
    "Sometimes I don't realise I was hungry until I ate something ",
  MINT_Sati_21 =
    "I don't always feel the need to drink until I am really thirsty",
  MINT_Olfa_22 = 
    "I often check the smell of my armpits",
  MINT_Olfa_23 = 
    "I often check the smell of my own breath",
  MINT_Olfa_24 = 
    "I often check the smell of my farts",
  MINT_Resp_25 = 
    "In general, I am very sensitive to changes in my breathing",
  MINT_Resp_26 = 
    "I can notice even very subtle changes in my breathing",
  MINT_Resp_27 =
    "I am always very aware of how I am breathing, even when I am calm",
  MINT_Card_28 = 
    "In general, I am very sensitive to changes in my heart rate",
  MINT_Card_29 = 
    "I often notice changes in my heart rate",
  MINT_Card_30 =
    "I can notice even very subtle changes in the way my heart beats",
  MINT_Gast_31 =
    "I can notice even very subtle changes in what my stomach is doing",
  MINT_Gast_32 = 
    "In general, I am very sensitive to what my stomach is doing",
  MINT_Gast_33 =
    "I am always very aware of what my stomach is doing, even when I am calm"
)
```

```{r}
cleanlabels <- function(x, qname=FALSE) {
  is_factor <- FALSE
  if(is.factor(x)) {
    is_factor <- TRUE
    vec <- x
    x <- levels(x)
  }
  
  x <- x |> 
    str_replace("MINT_", ifelse(qname, "MINT - ", "")) |> 
    str_replace("_", " - ") 
  
  if(is_factor) {
    levels(vec) <- x
    return(vec)
  }
  x
}
```


# 3. Replicate EGA from InteroceptionScale Study 2

```{r}
#| eval: false

ega <- bootEGA(
  mint_items,
  EGA.type = "hierEGA",
  model = "glasso",  # BGGM
  algorithm = "leiden",
  # consensus.method = "lowest_tefi",
  allow.singleton = TRUE,
  type="resampling",
  plot.itemStability=FALSE,
  typicalStructure=TRUE,
  plot.typicalGraph=FALSE,
  iter=500,
  seed=3, 
  ncores = 4)

save(ega, file="models/ega.RData")
```

```{r}
#| fig-width: 12
#| fig-height: 14
#| fig-keep: "first"

load("models/ega.RData")

EGAnet::dimensionStability(ega)
itemstability <- EGAnet::itemStability(ega, IS.plot=FALSE)
p_stab <- plot(itemstability)
p_stab

EGAnet::genTEFI(ega$EGA)


ega$EGA
```

Lower-order dimensions are very stable. Higher-order dimensions are mostly stable but a few items show ambiguity (Derm 16, 17 & 18, Olfa 22, 23 & 24).

Deficit metacluster is less consistent (0.546) - less robust, possibly due to heterogeneous items

Network characteristics:
Lower-order network - 33 nodes, 194 edges, edge density 0.367.
Most edges are small (0.071 ± 0.108), range -0.088 to 0.48 → mostly weak to moderate partial correlations.
Higher-order network - 11 nodes (representing lower-order communities), 36 edges, edge density 0.655 → more connected than lower-order network.
Edge weights are slightly stronger on average (0.096 ± 0.103), some negative → stronger inter-community relationships, but some inverse relationships.

Total Entropy Fit Index:
More negative TEFI = better fit of bifactor/hierarchical model.
Results suggest a bifactor structure fits better than just correlated lower-order factors.


# 4. Clustering

```{r}
# Scale data
mint_scaled <- as.data.frame(scale(mint_lowerorder))
```

## 4.1. Estimate Number of Clusters

```{r}
cluster_estimate <- n_clusters(
  mint_scaled,
  package = c("all"),
  n_max = 10,
)
cluster_estimate

plot(cluster_estimate)
```

Method Agreement Procedure:
The choice of 3 clusters is supported by 10 (34.48%) methods out of 29 (Hartigan, Scott, trcovw, Tracew, Friedman, Rubin, Ratkowsky, Ball, SDindex, Mixture (VVE)).

## 4.2. Kmeans

```{r}
# 3 clusters

rez_kmeans3 <- cluster_analysis(mint_scaled, n = 3, method = "kmeans")

rez_kmeans3

plot(rez_kmeans3)

predict(rez_kmeans3) # Get clusters
```

Cluster 2 (Low across most facets) - Globally low interoception
Especially low on:
Cardioception (-0.84)
Respiroception (-0.82)
Dermal Hypersensitivity (-0.58)

Cluster 3 - High awareness/relatively low confusion
High on: Awareness facets (SexS, ExAc, RelA) and Visceroception (Card, Resp, Gast)
But slightly low on deficit facets (CaCo, Urin)

Cluster 1 (Mixed pattern) - High confusion/dysregulated interoception
High on Deficit facets (CaCo = +0.94, Urin = +0.71)
Moderate/high elsewhere

## 4.3. Hierarchical Clustering

```{r}
rez_hclust <- cluster_analysis(mint_scaled, n = 3, method = "hclust")

rez_hclust

plot(rez_hclust) + theme_modern()

predict(rez_hclust)
```

## 4.4. Hierarchical K-Means

```{r}
rez_hkmeans <- cluster_analysis(mint_scaled, n = 3, method = "hkmeans")

rez_hkmeans

plot(rez_hkmeans) + theme_modern()
```

## 4.5. K-Medoids (PAM)

```{r}
rez_pam <- cluster_analysis(mint_scaled, n = 3, method = "pam")

rez_pam

plot(rez_pam) + theme_modern()
```

## 4.6. DBSCAN

```{r}
eps <- n_clusters_dbscan(mint_scaled, min_size = 0.1)
eps

plot(eps) # Find the 'elbow' of the curve: 4.25
```

```{r}
rez_dbscan <- cluster_analysis(mint_scaled, method = "dbscan", dbscan_eps = 4.25)

rez_dbscan

plot(rez_dbscan) + theme_modern()
```

Suggests the existence of 1 cluster.
The 2 clusters accounted for 1.47% of the total variance of the original data.

## 4.7. K-Medoids with estimation of number of clusters (pamk)

```{r}
rez_pamk <- cluster_analysis(mint_scaled, method = "pamk")

rez_pamk

plot(rez_pamk) + theme_modern()
```

The 2 clusters accounted for 17.50% of the total variance of the original data.

## 4.8. Mixture

```{r}
rez_mixture <- cluster_analysis(mint_scaled, method = "mixture")

rez_mixture

plot(rez_mixture) + theme_modern()
```

The 4 clusters accounted for 13.43% of the total variance of the original data.


# 5. Metaclustering

```{r}
list_of_results <- list(
  rez_kmeans3, rez_hclust, rez_hkmeans, rez_pam, rez_pamk, rez_mixture, rez_dbscan
)

probability_matrix <- cluster_meta(list_of_results)

# Plot the matrix as a reordered heatmap
heatmap <- stats::heatmap(probability_matrix,
  scale = "none",
  col = grDevices::hcl.colors(256, palette = "inferno")
)
heatmap
```


# 6. Extract participants from clusters

K-Medoids (PAM) clustering was chosen as the final clustering solution due to its similarity to the meta-clustering solution and its cluster stability.

```{r}
cluster_assignment <- predict(rez_pam)

mega_df <- mega_df |>
  mutate(
    Clusters = factor(
      cluster_assignment,
      levels = c(1, 2, 3),
      labels = c("Cluster 1", "Cluster 2", "Cluster 3")
    )
  )
```


# 7. Analysis

## 7.1. Age
```{r}
# Effect size: small = .01, medium = .06, large = .14 

aov_age <- aov(Age ~ Clusters, data = mega_df)
summary(aov_age)
eta_squared(aov_age)  # effect size
```

F(2, 1686) = 30.55, p < .001
Age explains 3% of the variance in cluster membership - small effect size

## 7.2. Depression & Anxiety (PHQ-4)

```{r}
# Depression

aov_dep_phq <- aov(PHQ4_Depression ~ Clusters, data = mega_df)
summary(aov_dep_phq)
eta_squared(aov_dep_phq)  # effect size


# Anxiety

aov_anx_phq <- aov(PHQ4_Anxiety ~ Clusters, data = mega_df)
summary(aov_anx_phq)
eta_squared(aov_anx_phq)  # effect size
```

Depression (PHQ-4): F(2, 1127) = 57.05, p < .001
Depression (PHQ-4) scores explain 9% of the variance in cluster membership - medium effect size

Anxiety (PHQ-4): F(2, 1127) = 79.34, p < .001
Anxiety (PHQ-4) scores explain 12% of the variance in cluster membership - medium effect size

## 7.3. Gender

```{r}
# Cramér's V (2x3 table, df =2): small = .07, medium = .21, large = .35

gender_df <- mega_df |> filter(Gender != "Other")

# Chi-square
chi_gend <- chisq.test(gender_df$Gender, gender_df$Clusters)
chi_gend

# Effect size (Cramér's V)
cramers_v(table(gender_df$Gender, gender_df$Clusters))

# Observed frequencies
chi_gend$observed

# Expected frequencies - check <5
chi_gend$expected


# Row percentages (% of each gender in each cluster)
prop.table(table(gender_df$Gender, gender_df$Clusters), margin = 1) * 100

# Column percentages (% of each cluster that is each gender)
prop.table(table(gender_df$Gender, gender_df$Clusters), margin = 2) * 100
```

χ²(2) = 48.87, p < .001
Small-medium association - Cramér's V = 0.17

## 7.4. Mental Health Conditions (Diagnosed)

### 7.4.1. Mood

```{r}
# Remove NAs
mood_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Mood))

# Chi-square
chi_mood <- chisq.test(mood_df$Disorders_Psychiatric_Mood, mood_df$Clusters)
chi_mood

# Effect size (Cramér's V)
cramers_v(table(mood_df$Disorders_Psychiatric_Mood, mood_df$Clusters))

# Observed frequencies
chi_mood$observed

# Expected frequencies - check <5
chi_mood$expected
```

χ²(2) = 35.35, p < .001
Medium association - Cramér's V = 0.20

### 7.4.2. Anxiety

```{r}
# Remove NAs
anxiety_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Anxiety))

# Chi-square
chi_anxiety <- chisq.test(anxiety_df$Disorders_Psychiatric_Anxiety, anxiety_df$Clusters)
chi_anxiety

# Effect size (Cramér's V)
cramers_v(table(anxiety_df$Disorders_Psychiatric_Anxiety, anxiety_df$Clusters))

# Observed frequencies
chi_anxiety$observed

# Expected frequencies - check <5
chi_anxiety$expected
```

χ²(2) = 23.77, p < .001
Small-medium association - Cramér's V = 0.16

### 7.4.3. Eating Disorders

```{r}
# Remove NAs
eating_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Eating))

# Chi-square
chi_eating <- chisq.test(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters)
chi_eating

# Effect size (Cramér's V)
cramers_v(table(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters))

# Observed frequencies
chi_eating$observed

# Expected frequencies - check <5
chi_eating$expected
```

χ²(2) = 10.19, p = .006
Small association - Cramér's V = 0.10
However, 1 expected cell was below 5 (Cluster 2, True, 4.83)

```{r}
fisher.test(table(eating_df$Disorders_Psychiatric_Eating, eating_df$Clusters), 
            simulate.p.value = TRUE)
```

Due to small expected cell frequencies, Fisher's exact test was used in place of chi-square. There was a significant association between eating disorder diagnosis and cluster membership (p = .009), though the effect was small (Cramér's V = 0.10). Eating disorders were disproportionately represented in Cluster 3.

### 7.4.4 Autism

```{r}
# Remove NAs
autism_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_Autism))

# Chi-square
chi_autism <- chisq.test(autism_df$Disorders_Psychiatric_Autism, autism_df$Clusters)
chi_autism

# Effect size (Cramér's V)
cramers_v(table(autism_df$Disorders_Psychiatric_Autism, autism_df$Clusters))

# Observed frequencies
chi_autism$observed

# Expected frequencies - check <5
chi_autism$expected
```

χ²(2) = 21.53, p < .001
Small-medium association - Cramér's V = 0.15

### 7.4.5. ADHD

```{r}
# Remove NAs
adhd_df <- mega_df |> 
  filter(!is.na(Disorders_Psychiatric_ADHD))

# Chi-square
chi_adhd <- chisq.test(adhd_df$Disorders_Psychiatric_ADHD, adhd_df$Clusters)
chi_adhd

# Effect size (Cramér's V)
cramers_v(table(adhd_df$Disorders_Psychiatric_ADHD, adhd_df$Clusters))

# Observed frequencies
chi_adhd$observed

# Expected frequencies - check for <5
chi_adhd$expected
```

χ²(2) = 13.07, p = .001
Small association - Cramér's V = 0.12

## 7.5. Somatic Health Conditions (Diagnosed)

### 7.5.1. Afferent Sensitivity

```{r}
# Remove NAs
as_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_AfferentSensitivity))

# Chi-square
chi_as <- chisq.test(as_df$Disorders_Somatic_AfferentSensitivity, as_df$Clusters)
chi_as

# Effect size (Cramér's V)
cramers_v(table(as_df$Disorders_Somatic_AfferentSensitivity, as_df$Clusters))

# Observed frequencies
chi_as$observed

# Expected frequencies - check for <5
chi_as$expected
```

χ²(2) = 19.39, p < .001
Small-medium association - Cramér's V = 0.15

### 7.5.2. Central Sensitisation

```{r}
# Remove NAs
cs_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_CentralSensitization))

# Chi-square
chi_cs <- chisq.test(cs_df$Disorders_Somatic_CentralSensitization, cs_df$Clusters)
chi_cs

# Effect size (Cramér's V)
cramers_v(table(cs_df$Disorders_Somatic_CentralSensitization, cs_df$Clusters))

# Observed frequencies
chi_cs$observed

# Expected frequencies - check for <5
chi_cs$expected
```

χ²(2) = 12.92, p = .002
Small association - Cramér's V = 0.12

### 7.5.3. Autonomic Dysfunction

```{r}
# Remove NAs
ad_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_AutonomicDysfunction))

# Chi-square
chi_ad <- chisq.test(ad_df$Disorders_Somatic_AutonomicDysfunction, ad_df$Clusters)
chi_ad

# Effect size (Cramér's V)
cramers_v(table(ad_df$Disorders_Somatic_AutonomicDysfunction, ad_df$Clusters))

# Observed frequencies
chi_ad$observed

# Expected frequencies - check for <5
chi_ad$expected
```

χ²(2) = 6.70, p = .035
Small association - Cramér's V = 0.08

### 7.5.4. Immune-Inflammatory

```{r}
# Remove NAs
ii_df <- mega_df |> 
  filter(!is.na(Disorders_Somatic_ImmuneInflammatory))

# Chi-square
chi_ii <- chisq.test(ii_df$Disorders_Somatic_ImmuneInflammatory, ii_df$Clusters)
chi_ii

# Effect size (Cramér's V)
cramers_v(table(ii_df$Disorders_Somatic_ImmuneInflammatory, ii_df$Clusters))

# Observed frequencies
chi_ii$observed

# Expected frequencies - check for <5
chi_ii$expected
```

χ²(2) = 22.81, p < .001
Small-medium association - Cramér's V = 0.16


# 8. Multinomial Logistic Regression

```{r}
# Set a reference cluster - what should this be?
mega_df$Clusters_relev <- relevel(mega_df$Clusters, ref = "Cluster 1")

# Make dummy variable
mega_df$Male <- rec(mega_df$Gender, rec = "Male = 1; Female = 0; else = NA")

multinom <- multinom(
  Clusters_relev ~ Age + PHQ4_Depression + PHQ4_Anxiety + Gender + Disorders_Psychiatric_Mood + Disorders_Psychiatric_Anxiety + Disorders_Psychiatric_Eating + Disorders_Psychiatric_Autism + Disorders_Psychiatric_ADHD + Disorders_Somatic_AfferentSensitivity + Disorders_Somatic_CentralSensitization + Disorders_Somatic_AutonomicDysfunction + Disorders_Somatic_ImmuneInflammatory,
  data = mega_df
)

summary(multinom)

tidy(multinom, exponentiate = TRUE, conf.int = TRUE)  # odds ratios
```

For Cluster 2, higher anxiety symptoms significantly increased the odds of membership (OR = 1.26, p = .014), as did the presence of immune-inflammatory somatic disorders (OR = 1.64, p = .027). No other predictors significantly distinguished Cluster 2 from Cluster 1.
For Cluster 3, younger age was associated with greater odds of membership (OR = 0.97, p < .001). Higher anxiety symptoms also predicted Cluster 3 membership (OR = 1.29, p = .002), and men were less likely than women to belong to this cluster (OR = 0.45, p < .001). Several additional predictors showed marginal associations (ps = .050–.073) but were not significant.
